### 2021-08-26 scripts/FINAL PROCESSING/Analysis/Final analysis.Rmd
data

---
title: "The Role of Theory of Mind and Interoception in Deception and Psychopathy "
subtitle: "Statistical Analysis"
output:
  html_document:
    theme: journal
    footer: LIE questionnaire validation
    highlight: pygments
    toc: yes
    toc_depth: 4
    toc_float: yes
    df_print: kable
    code_folding: hide
  word_document: default
  pdf_document:
    toc: yes
    toc_depth: '4'
editor_options: 
  chunk_output_type: console
---

Final Year Project by Low Jia Rong (U1730052L)

# Packages and Setup

## Libraries
```{r setup, include=TRUE, warning = FALSE, message = FALSE}
setwd("~/Dropbox/FYP_Jia Rong/R Processing/FINAL PROCESSING/Analysis")
library(tidyverse)
library(lme4)
library(ggplot2)
library(glmmTMB)
library(correlation)
library(sjPlot)
#library(easystats)

library(report)
library(correlation)
library(parameters)
library(see)

descriptive_statistics <- function(df, begins_with){
  df %>% 
    select(dplyr::starts_with(begins_with)) %>% 
    report() %>% 
    as.data.frame() %>% 
    select(-one_of(c("n_Obs", "Median", "MAD", "n_Missing"))) %>% 
    print()

  plot(df %>% 
    select(dplyr::starts_with(begins_with)) %>% 
    bayestestR::estimate_density(method = "KernSmooth") %>% 
    plot() + 
      xlab("Score") + ylab("Frequency") +
    see::theme_modern())
}

```

## Read data
```{r read_data, include=TRUE, warning = FALSE, message = FALSE}
Behav <- read.csv('BehaviouraldataNEW.csv', stringsAsFactors = FALSE, header = TRUE)[-c(1)] %>% 
  rename(participant = ID,
         truthfulness = condition) %>% 
  mutate(ConditionSP = rep(c("social","polygraph"), each = 40, times = 30))

Behav[stringr::str_detect(names(Behav), "subresp")] <- effectsize::change_scale(Behav[stringr::str_detect(names(Behav), "subresp")], from = c(0, 1), to = c(10, -10))

Phy <- read.csv('DT_04.csv', stringsAsFactors = FALSE, header = TRUE) [-c(1),] %>% 
  rename(participant = Participant.number,) %>% 
  mutate(Trial = rep(1:80, each = 4, times = 30))

data <- merge(Behav, Phy, by = c("participant" ,"Trial"), all = TRUE) %>% 
  arrange(participant,Trial) %>% 
  mutate(participant = as.factor(participant),
         ConditionSP = as.factor(ConditionSP))

Demo <- read.csv('demo_30.csv', stringsAsFactors = FALSE, header = TRUE)[-c(2,3,4:6,9,11,12,18)]

```

```{r , include=FALSE, warning = FALSE, message = FALSE}
## Cleaning data
data <- data %>% 
  filter(Inclusion != "0")

```

# Descriptive Statistics
## Participants Demographics Information
Summary of participants Gender, Age, and Handedness
```{r , include=TRUE, warning = FALSE, message = FALSE}

#General Descriptive Stats
report(Demo)
display(Demo)

#Gender
Gender <- Demo %>% 
  group_by(Gender) %>% 
  summarize(n = n(),
            Ratio = n() / nrow(Demo))
report(Demo$Gender)

as.data.frame(table(Demo$Gender)) %>%
  ggplot(aes(x="", y =Freq, fill = reorder(Var1, -Freq))) +
  geom_bar(width = 1, stat = "identity") +
  labs(fill = "Gender") +
  coord_polar("y", start = 0, direction = -1) +
  scale_fill_brewer(palette="Blues") +
  theme_void() +
  theme(legend.text = element_text(size = 12)) +
  theme(legend.title = element_text(face = "bold", size = 20))

#Age
Age <- Demo [-c(1),] %>% 
  summarise(mean(Age),sd(Age))
report(Demo$Age)

as.data.frame(table(Demo$Age)) %>%
  ggplot(aes(x="", y =Freq, fill = reorder(Var1, -Freq))) +
  geom_bar(width = 1, stat = "identity") +
  labs(fill = "Age") +
  coord_polar("y", start = 0, direction = 1) +
  scale_fill_brewer(palette="Oranges", breaks=c("18","19","20","21","22","24","25")) +
  theme_void() +
  theme(legend.text = element_text(size = 15)) +
  theme(legend.title = element_text(face = "bold", size = 20))

#Handedness
Handedness <- Demo %>% 
  group_by(handedness) %>% 
  summarize(n = n(),
            Ratio = n() / nrow(Demo))
report(Demo$handedness)

as.data.frame(table(Demo$handedness)) %>%
  ggplot(aes(x="", y =Freq, fill = reorder(Var1, Freq))) +
  geom_bar(width = 1, stat = "identity") +
  labs(fill = "Handedness") +
  coord_polar("y", start = 0, direction = -1) +
  scale_fill_brewer(palette="greens") +
  theme_void() +
  theme(legend.text = element_text(size = 12)) +
  theme(legend.title = element_text(face = "bold", size = 20))

#Ethnicity
as.data.frame(table(Demo$Ethnicity)) %>%
  ggplot(aes(x="", y =Freq, fill = reorder(Var1, -Freq))) +
  geom_bar(width = 1, stat = "identity") +
  labs(fill = "Ethnicity") +
  coord_polar("y", start = 0, direction = -1) +
  scale_fill_brewer(palette="Pastel1") +
  theme_void() +
  theme(legend.text = element_text(size = 12)) +
  theme(legend.title = element_text(face = "bold", size = 20))

#Religion
as.data.frame(table(Demo$Religion)) %>%
  ggplot(aes(x="", y =Freq, fill = reorder(Var1, -Freq))) +
  geom_bar(width = 1, stat = "identity") +
  labs(fill = "Religion") +
  coord_polar("y", start = 0, direction = -1) +
  scale_fill_brewer(palette="Set2") +
  theme_void() +
  theme(legend.text = element_text(size = 12)) +
  theme(legend.title = element_text(face = "bold", size = 20))

#Nationality
as.data.frame(table(Demo$Nationality)) %>%
  ggplot(aes(x="", y =Freq, fill = reorder(Var1, -Freq))) +
  geom_bar(width = 1, stat = "identity") +
  labs(fill = "Nationality") +
  coord_polar("y", start = 0, direction = -1) +
  scale_fill_brewer(palette="Set2") +
  theme_void() +
  theme(legend.text = element_text(size = 12)) +
  theme(legend.title = element_text(face = "bold", size = 20))

```

## Tasks Descriptive Statistics
Summary of Questionnaires and  Behavioural Tasks Data
```{r , include=TRUE, warning = FALSE, message = FALSE}

des_stat <- data %>% 
group_by(participant) %>% 
  summarise_all(mean) %>% 
  mutate(yoni_new = yoni_affective + yoni_cognitive) %>% 
  select(-ends_with("Total"),
          -ends_with("TotalAvg"))

#Questionnaires
descriptive_statistics(des_stat, "BES_")          #Basic Empathy/ToM
descriptive_statistics(des_stat, "MAIA")          #Interoceptive Awareness
descriptive_statistics(des_stat, "SRP3_")         #Psychopathic Traits

#Behavioural Tasks
descriptive_statistics(des_stat, "HCT") +
  facet_wrap(~Parameter, scales="free")           #Heartbeat Counting Task (Interoception)
descriptive_statistics(des_stat, "yoni_")         #Yoni Task (ToM)

#Deception Task Subjective Response Data
df_sub <- Behav %>% 
  filter(truthfulness == "LIE") %>% 
  group_by(participant, ConditionSP) %>% 
  summarise_all(mean) %>% 
  select(1,2,9) %>% 
  pivot_wider(names_from = ConditionSP, values_from = subresp) %>% 
  rename(sub_poly = polygraph,
         sub_soc = social) %>% 
  mutate(participant = as.factor(participant)) %>% 
  mutate(sub_total = (sub_poly + sub_soc)/2) %>% 
  ungroup(participant)

descriptive_statistics(df_sub, "sub_")          #Subjective Response

```

## Participants Individual Summary Results for Behavioural Tasks (Deception, HCT & Yoni Task)
```{r , include=TRUE, warning = FALSE, message = FALSE}

df_descriptive <- data %>% 
  filter(truthfulness != "TRUTH") %>% 
  group_by(participant, ConditionSP) %>% 
  summarise_all(mean)

df_descriptive <- df_descriptive %>% 
  select(c(1,2,9,36:42)) %>% 
  pivot_wider(names_from = ConditionSP, values_from = subresp) %>% 
  rename(PolyCon_subresp = polygraph,
         SociCon_subresp = social)
display(df_descriptive)

```

```{r , include=FALSE, warning = FALSE, message = FALSE}

## Plot of Average Subjective Response of Lies for Each Participant (REMOVE)
plot_desriptive <- data %>% 
  filter(truthfulness != "TRUTH") %>% 
  group_by(participant, ConditionSP) %>% 
  summarise_all(mean) %>% 
  select(c(1,2,9)) %>% 
  filter(participant != "3",
         participant != "15",
         participant != "19",
         participant != "23")

pd <- position_dodge(0.1)
plot_desriptive %>% 
  mutate(participant = as.factor(participant)) %>% 
  ggplot(aes(x = ConditionSP, y = subresp, color = participant, group=participant)) +
  geom_point(position = pd) +
  geom_line(position = pd) +
  xlab("Condition") + ylab("Subjective Response")+
  ggtitle("Plot of Average Subjective Responses for Lies")  + 
  see::theme_modern() +
  theme(plot.title = element_text(size=12))

```

## Summary of Tasks
average and SD of each task
```{r , include=TRUE, warning = FALSE, message = FALSE}
Task_summary <- data.frame(
  Mean = sapply(df_descriptive[2:ncol(df_descriptive)], mean, na.rm = TRUE),
  SD = sapply(df_descriptive[2:ncol(df_descriptive)], sd, na.rm = TRUE)
)
print(Task_summary)

```

# Correlation Matrix of all Variables
```{r , include=TRUE, warning = FALSE, message = FALSE}

df_correlation <- data %>% 
  filter(truthfulness != "TRUTH") %>% 
  filter(Condition == "R") %>% 
  select(c(1,8,50,55:90)) %>% 
  mutate(participant = as.factor(participant),
         ConditionSP = as.factor(ConditionSP))

Corr_Matrix <- correlation::correlation(
  select(df_correlation,
         participant, 
         ConditionSP, 
         subresp, 
         starts_with("ECG"), 
         starts_with("RSP"),
         starts_with("EDA")),
  multilevel = F, p_adjust = "none")

dplyr::filter(Corr_Matrix, Parameter1 == "subresp") %>% 
  arrange(desc(abs(r)))

```

# Hypothesis Analysis
## Subjective Response Comparison
Linear Mixed Model Regression

### Hypothesis 1A: Predicted Values of Subjective Response by Yoni Task
```{r , include=TRUE, warning = FALSE, message = FALSE}

df <- data %>% 
  mutate(participant = as.factor(participant),
         ConditionSP = as.factor(ConditionSP)) %>% 
  filter(truthfulness == "LIE") %>% 
  filter(Inclusion2 != "0") %>% 
  filter(participant != "3",
         participant != "15",
         participant != "19",
         participant != "23") %>% 
  filter(Condition == "Q") %>% 
  mutate(yoni_new = yoni_affective + yoni_cognitive) 
df$ConditionSP <- relevel(df$ConditionSP, ref = "social")

# TOM Hypothesis 1A: SUBRESP -> YONI
model_sub1 <- lme4::lmer(subresp ~ ConditionSP * yoni_new + (1|participant), data = df)  #explorative
parameters::parameters(model_sub1)
plot_model(model_sub1, type = "pred", terms = c("yoni_new", "ConditionSP"), axis.title = c("Yoni Task", "Subjective Response"), title = "Predicted Values of Subjective Response by Yoni Task", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_sub1)

```

### Hypothesis 1A: Predicted Values of Subjective Response by BES Questionnaire
```{r , include=TRUE, warning = FALSE, message = FALSE}
# TOM Hypothesis 1A: SUBRESP -> YONI/BES
model_sub2 <- lme4::lmer(subresp ~ ConditionSP * BES_Total + (1|participant), data = df)
parameters::parameters(model_sub2)
plot_model(model_sub2, type = "pred", terms = c("BES_Total", "ConditionSP"), axis.title = c("BES Questionnaire", "Subjective Response"), title = "Predicted Values of Subjective Response by BES Questionnaire", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_sub2)
```

### Hypothesis 1C: Predicted Values of Subjective Response by SRP3
```{r , include=TRUE, warning = FALSE, message = FALSE}
#TOM Hypothesis 1C: SUBRESP -> SRP3
model_sub3 <- lme4::lmer(subresp ~ ConditionSP * SRP3_Total + (1|participant), data = df)
parameters::parameters(model_sub3)
plot_model(model_sub3, type = "pred", terms = c("SRP3_Total", "ConditionSP"), axis.title = c("SRP3 Inventory", "Subjective Response"), title = "Predicted Values of Subjective Response by SRP3 Inventory", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_sub3)
```

### Hypothesis 2A: Predicted Values of Subjective Response by HCT Accuracy
```{r , include=TRUE, warning = FALSE, message = FALSE}
df$ConditionSP <- relevel(df$ConditionSP, ref = "polygraph")

#Interoception Hypothesis 2A: SUBRESP -> HCT
model_sub4 <- lme4::lmer(subresp ~ ConditionSP * HCT_Accuracy + (1|participant), data = df)
parameters::parameters(model_sub4)
plot_model(model_sub4, type = "pred", terms = c("HCT_Accuracy", "ConditionSP"), axis.title = c("HCT Accuracy", "Subjective Response"), title = "Predicted Values of Subjective Response by HCT Accuracy", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_sub4)
```

### Hypothesis 2A: Predicted Values of Subjective Response by MAIA2
```{r , include=TRUE, warning = FALSE, message = FALSE}
#Interoception Hypothesis 2A: SUBRESP -> MAIA2
model_sub5 <- lme4::lmer(subresp ~ ConditionSP * MAIA_Total + (1|participant), data = df)
parameters::parameters(model_sub5)
plot_model(model_sub5, type = "pred", terms = c("MAIA_Total", "ConditionSP"), axis.title = c("MAIA2", "Subjective Response"), title = "Predicted Values of Subjective Response by MAIA2", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_sub5)
```

### Hypothesis 2C: Predicted Values of Subjective Response by Primary SRP3 Subfactor
```{r , include=TRUE, warning = FALSE, message = FALSE}
#Interoception Hypothesis 2C: SUBRESP -> SRP3_PRI
model_sub6 <- lme4::lmer(subresp ~ ConditionSP * SRP3_PRI + (1|participant), data = df)
parameters::parameters(model_sub6)
plot_model(model_sub6, type = "pred", terms = c("SRP3_PRI", "ConditionSP"), axis.title = c("SRP3 (PRI) Inventory", "Subjective Response"), title = "Predicted Values of Subjective Response by SRP3 (PRI)", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_sub6)

```

## ECG Physio scores comparison
Linear Mixed Model Regression

### Hypothesis 1A: Predicted Values of Heartrate Acceleration by Yoni Task
```{r , include=TRUE, warning = FALSE, message = FALSE}

df2 <- data %>% 
  mutate(participant = as.factor(participant),
         ConditionSP = as.factor(ConditionSP)) %>% 
  filter(truthfulness == "LIE") %>% 
  filter(Inclusion2 != "0") %>%
#  filter(participant != "22") %>% 
#  filter(participant != "23") %>%         
  filter(participant != "30") %>% 
  filter(participant != "15") %>% 
  filter(Condition == "Q") %>% 
  mutate(yoni_new = yoni_affective + yoni_cognitive) 
df2$ConditionSP <- relevel(df2$ConditionSP, ref = "social")

#ECG_RATE_MAX
#TOM Hypothesis 1A: ECG_Rate_Max -> YONI
model_ecg1 <- lme4::lmer(ECG_Rate_Max ~ ConditionSP * yoni_new + (1|participant), data = df2)  #explorative
parameters::parameters(model_ecg1)
plot_model(model_ecg1, type = "pred", terms = c("yoni_new", "ConditionSP"), axis.title = c("Yoni Task", "Heartrate Acceleration"), title = "Predicted Values of Heartrate Acceleration by Yoni Task ", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_ecg1)
```

### Hypothesis 1A: Predicted Values of Heartrate Acceleration by BES Questionniare
```{r , include=TRUE, warning = FALSE, message = FALSE}
#TOM Hypothesis 1A: ECG_Rate_Max -> BES
model_ecg2 <- lme4::lmer(ECG_Rate_Max ~ ConditionSP * BES_Total + (1|participant), data = df2)
parameters::parameters(model_ecg2)
plot_model(model_ecg2, type = "pred", terms = c("BES_Total", "ConditionSP"), axis.title = c("BES Questionnaire", "Heartrate Acceleration"), title = "Predicted Values of Heartrate Acceleration by BES Questionnaire", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_ecg2)
```

### Hypothesis 1C: Predicted Values of Heartrate Acceleration by SRP3
```{r , include=TRUE, warning = FALSE, message = FALSE}
#TOM Hypothesis 1C: ECG_Rate_Max -> SRP3
model_ecg3 <- lme4::lmer(ECG_Rate_Max ~ ConditionSP * SRP3_Total + (1|participant), data = df2)
parameters::parameters(model_ecg3)
plot_model(model_ecg3, type = "pred", terms = c("SRP3_Total", "ConditionSP"), axis.title = c("SRP3 Inventory", "Heartrate Acceleration"), title = "Predicted Values of Heartrate Acceleration by SRP3 Inventory", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_ecg3)
```

### Hypothesis 2A: Predicted Values of Heartrate Acceleration by HCT Accuracy
```{r , include=TRUE, warning = FALSE, message = FALSE}
df2$ConditionSP <- relevel(df2$ConditionSP, ref = "social")

#Interoception Hypothesis 2A: ECG_Rate_Max -> HCT
model_ecg4 <- lme4::lmer(ECG_Rate_Max ~ ConditionSP * HCT_Accuracy + (1|participant), data = df2)
parameters::parameters(model_ecg4)
plot_model(model_ecg4, type = "pred", terms = c("HCT_Accuracy", "ConditionSP"), axis.title = c("HCT Accuracy", "Heartrate Acceleration"), title = "Predicted Values of Heartrate Acceleration by HCT Accuracy", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_ecg4)
```

### Hypothesis 2A: Predicted Values of Heartrate Acceleration by MAIA2
```{r , include=TRUE, warning = FALSE, message = FALSE}
#Interoception Hypothesis 2A: ECG_Rate_Max -> MAIA2
model_ecg5 <- lme4::lmer(ECG_Rate_Max ~ ConditionSP * MAIA_Total + (1|participant), data = df2)
parameters::parameters(model_ecg5)
plot_model(model_ecg5, type = "pred", terms = c("MAIA_Total", "ConditionSP"), axis.title = c("MAIA2", "Heartrate Acceleration"), title = "Predicted Values of Heartrate Acceleration by MAIA2", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_ecg5)
```

### Hypothesis 2C: Predicted Values of Heartrate Acceleration by Primary SRP3 Subfactor
```{r , include=TRUE, warning = FALSE, message = FALSE}
#Interoception Hypothesis 2C: ECG_Rate_Max -> SRP3_PRI
model_ecg6 <- lme4::lmer(ECG_Rate_Max ~ ConditionSP * SRP3_PRI + (1|participant), data = df2)
parameters::parameters(model_ecg6)
plot_model(model_ecg6, type = "pred", terms = c("SRP3_PRI", "ConditionSP"), axis.title = c("SRP3 (PRI) Inventory", "Heartrate Acceleration"), title = "Predicted Values of Heartrate Acceleration by SRP3 (PRI) ", legend.title = "Condition") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_ecg6)
```

## Correalation Analysis
### Hypothesis 1B: Correlation Between ToM and Psychopathy
```{r , include=TRUE, warning = FALSE, message = FALSE}
df5 <- data %>% 
  mutate(yoni_new = yoni_affective + yoni_cognitive) %>% 
group_by(participant) %>% 
  summarise_all(mean)

#Hypothesis 1b
##Yoni Task (1b)
model_H1 <- lm(yoni_affective ~  SRP3_Total, data = df5)
parameters::parameters(model_H1)
plot_model(model_H1, type = "pred", terms = c("SRP3_Total"), axis.title = c("SRP3 Inventory", "Yoni Task (Affective)"), title = "Relationship Between Yoni Task (Affective) and SRP3") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_H1)

correlation::correlation(select(df5,
         yoni_affective,
         SRP3_Total), method = "pearson")

##Basic Empathy Scale (1b)
model_H1a <- lm(BES_Affective ~  SRP3_Total, data = df5)
parameters::parameters(model_H1a)
plot_model(model_H1a, type = "pred", terms = ("SRP3_Total"),  axis.title = c("SRP3 Inventory", "BES (Affective)"), title = "Relationship Between BES (Affective) Score and SRP3") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_H1a)

correlation::correlation(select(df5,
         BES_Affective,
         SRP3_Total), method = "pearson")

```

### Hypothesis 2B: Correlation Between Interoception and Primary Psychopathy
```{r , include=TRUE, warning = FALSE, message = FALSE}
#Hypothesis 2b
##Heartbeat Counting Task(2b)
model_H2 <- lm(HCT_Accuracy ~  SRP3_PRI, data = df5)
parameters::parameters(model_H2)
plot_model(model_H2, type = "pred", terms = ("SRP3_PRI"), axis.title = c("SRP3 (PRI) Inventory", "HCT Accuracy"), title = "Relationship Between HCT Accuracy and SRP3 (PRI)") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_H2)

correlation::correlation(select(df5,
         HCT_Accuracy,
         SRP3_PRI), method = "pearson")

##MAIA2 (2b)
model_H2a <- lm(MAIA_Total ~  SRP3_PRI, data = df5)
parameters::parameters(model_H2)
plot_model(model_H2, type = "pred", terms = ("SRP3_PRI"), axis.title = c("SRP3 (PRI) Inventory", "MAIA2"), title = "Relationship Between MAIA2 and SRP3 (PRI)") + theme_sjplot(base_size = 10, base_family = "Times New Roman")
report(model_H2a)

correlation::correlation(select(df5,
         MAIA_Total,
         SRP3_PRI), method = "pearson")
```

### Correlation Between Yoni Task and Basic Empathy Scale
```{r , include=TRUE, warning = FALSE, message = FALSE}
#Correlation between Yoni Task and BES Scale
correlation::correlation(select(df5,
         yoni_new,
         BES_Total), method = "pearson")

correlation::correlation(select(df5,
         yoni_affective,
         BES_Affective), method = "pearson")
```