### 2021-09-10 1_descriptive.Rmd
started on descriptive

---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Part 1

Here's a cool plot:

```{r 1_plot_scatter_basic, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi=150}
ggplot(df, aes(x=V1, y=V2, color=Participant)) + 
  geom_point() +
  see::theme_modern()
```

## Part 2

That's another great plot:

```{r 1_plot_density, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi=150}
plot(bayestestR::estimate_density(df[c("V1", "V2")])) +
  see::theme_blackboard()
```

## Part 3

```{r}
ggplot(df, aes(yoni_total)) + geom_histogram()
ggplot(df, aes(BES_total)) + geom_histogram()
ggplot(df, aes(yoni_total, BES_total)) + geom_point()
ggplot(df, aes(HCT_accuracy)) + geom_histogram()
ggplot(df, aes(MAIA_total)) + geom_histogram()
ggplot(df, aes(HCT_accuracy, MAIA_total)) + geom_point()
```