### 2021-09-27 1_descriptive.Rmd
lmer and descriptive (dt)

---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Part 1

Here's a cool plot:

```{r 1_plot_scatter_basic, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi=150}
ggplot(df, aes(x=V1, y=V2, color=Participant)) + 
  geom_point() +
  see::theme_modern()
```

## Part 2

That's another great plot:

```{r 1_plot_density, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi=150}
plot(bayestestR::estimate_density(df[c("V1", "V2")])) +
  see::theme_blackboard()
```

## Part 3

```{r independent var}
ggplot(df, aes(yoni_total)) + geom_histogram()
ggplot(df, aes(BES_total)) + geom_histogram()
ggplot(df, aes(yoni_total, BES_total)) + geom_point()
ggplot(df, aes(HCT_accuracy)) + geom_histogram()
ggplot(df, aes(MAIA_total)) + geom_histogram()
ggplot(df, aes(HCT_accuracy, MAIA_total)) + geom_point()
```

```{r dependent var}
ggplot(df, aes(instruction, DT_confidence)) + geom_boxplot()
ggplot(df, aes(instruction, DT_confidenceRT)) + geom_boxplot()
ggplot(df, aes(condition, DT_confidence)) + geom_boxplot()
ggplot(df, aes(condition, DT_confidenceRT)) + geom_boxplot()
```

```{r}
df.social <- df %>% filter(condition == 'Social')
ggplot(df.social, aes(yoni_total, DT_confidence_avg)) + geom_point()
ggplot(df.social, aes(BES_total, DT_confidence_avg)) + geom_point()

df.polyg <- df %>% filter(condition == 'Polygraph')
ggplot(df.polyg, aes(HCT_accuracy, DT_confidence_avg)) + geom_point()
ggplot(df.polyg, aes(MAIA_total, DT_confidence_avg)) + geom_point()
```