### 2021-09-27 2_inferential.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---

Here is another analysis that is contained in a separate file. Let's check-out this linear regression model (note that, in the code chunk parameters, I multiplied `figheight` by 2 to have a taller plot):

```{r 2_plot_model, warning=FALSE, message=FALSE, fig.heiht=figheight * 2, fig.width=figwidth, dpi=150}
model <- lm(Petal.Length ~ Sepal.Length, data=iris)
performance::check_model(model)
```

```{r}
y.var <- c('DT_confidence', 'DT_confidence', 'DT_confidenceRT')
x.var <- c('DT_confidenceRT', 'condition*instruction', 'condition*instruction')

for (i in seq_along(y.var)){
  for (j in seq_along(x.var)){
    if (i==j){
      model <- lmer(paste(y.var[i],' ~', x.var[j],' + (1|ID) + (1|item) + (1|trial)'), df)
      print(report(model))
      print(parameters(model, effects = "fixed"))
      if (grepl('\\*', x.var[j]) == TRUE){
        p <- interactions::cat_plot(model, pred = instruction, modx = condition)
      }
      else{
        viz_data <- estimate_relation(model)
        p <- ggplot(data = viz_data, aes(x = DT_confidenceRT, y = Predicted)) +
          geom_line() +
          geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3) +
          geom_point(data = df, aes(x = DT_confidenceRT, y = DT_confidence, color = ID))
      }
      print(p)
    }
  }
}
```

```{r}
model <- lmer(DT_confidence ~ DT_confidenceRT + (1|ID) + (1|item) + (1|trial), df)
report(model)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model)
p <- ggplot(data = viz_data, aes(x = DT_confidenceRT, y = Predicted)) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3) +
  geom_point(data = df, aes(x = DT_confidenceRT, y = DT_confidence, color = ID))
p
#R2 does not account for random effects
#weirdly significant given the plot
#what's the include_random thing (6.5 vs 6.7)
```

```{r}
model <- lmer(DT_confidence ~ condition*instruction + (1|ID) + (1|item) + (1|trial), df)
report(model)
parameters(model, effects = "fixed")
interactions::cat_plot(model, pred = instruction, modx = condition)
#R2 does not account for random effects
#interaction is significant
```

```{r}
model <- lmer(DT_confidenceRT ~ condition*instruction + (1|ID) + (1|item) + (1|trial), df)
#actually this is not the reaction time? it's the reaction time to the question of how confident
```