### 2021-10-07 1_descriptive.Rmd
descriptive and interaction models
condition*instruction
style*instruction

---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Part 1

Here's a cool plot:

```{r 1_plot_scatter_basic, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi=150}
ggplot(df, aes(x=V1, y=V2, color=Participant)) + 
  geom_point() +
  see::theme_modern()
```

## Part 2

That's another great plot:

```{r 1_plot_density, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi=150}
plot(bayestestR::estimate_density(df[c("V1", "V2")])) +
  see::theme_blackboard()
```

## Part 3

```{r}
df.plot <- df %>% 
  group_by(ID, condition, instruction) %>% 
  summarise(confidence = mean(DT_confidence, na.rm = TRUE),
            RT = mean(DT_RT, na.rm = TRUE),
            yoni_total = mean(yoni_total, na.rm = TRUE),
            BES_total = mean(BES_total, na.rm = TRUE),
            HCT_accuracy = mean(HCT_accuracy, na.rm = TRUE),
            MAIA_total = mean(MAIA_total, na.rm = TRUE))
```

```{r independent var}
ggplot(df.plot, aes(yoni_total)) + geom_histogram()
ggplot(df.plot, aes(BES_total)) + geom_histogram()
ggplot(df.plot, aes(yoni_total, BES_total)) + geom_point()
ggplot(df.plot, aes(HCT_accuracy)) + geom_histogram()
ggplot(df.plot, aes(MAIA_total)) + geom_histogram()
ggplot(df.plot, aes(HCT_accuracy, MAIA_total)) + geom_point()
```

```{r dependent var}
ggplot(df, aes(DT_confidence)) + geom_density()
ggplot(df, aes(DT_RT)) + geom_density()

ggplot(df, aes(instruction, DT_confidence)) + geom_boxplot()
# much higher confidence in truth
ggplot(df, aes(instruction, DT_RT)) + geom_boxplot()
# no diff in reaction time in truth or lie

ggplot(df, aes(condition, DT_confidence)) + geom_boxplot()
ggplot(df, aes(condition, DT_RT)) + geom_boxplot()
# not much diff for diff conditions

ggplot(df, aes(DT_RT, DT_confidence)) + geom_point()
# RT and confidence don't seem to have a correlation
```

```{r both}
# is there a way to loop this
ggplot(df.plot, aes(x = yoni_total, y = confidence, colour = instruction)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(~condition)

ggplot(df.plot, aes(x = BES_total, y = confidence, colour = instruction)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(~condition)
# both yoni and BES do not seem to have effect in social condition, higher yoni/BES increases confidence in truth and decreases confidences in lie

ggplot(df.plot, aes(x = HCT_accuracy, y = confidence, colour = instruction)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(~condition)
# same effect in both social and polygraph, higher HCT accuracy decreases confidence in truth and increases confidence in lie

ggplot(df.plot, aes(x = MAIA_total, y = confidence, colour = instruction)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(~condition)
# same effect for lie in both social and polygraph, higher MAIA increases confidence
# higher MAIA increases confidence in lie for social and not polygraph

# interesting that the social tests are more correlated with polygraph and vice versa
```