### 2021-10-07 2_inferential.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---

Here is another analysis that is contained in a separate file. Let's check-out this linear regression model (note that, in the code chunk parameters, I multiplied `figheight` by 2 to have a taller plot):

```{r 2_plot_model, warning=FALSE, message=FALSE, fig.heiht=figheight * 2, fig.width=figwidth, dpi=150}
model <- lm(Petal.Length ~ Sepal.Length, data=iris)
performance::check_model(model)
```

```{r loop archive}
y.var <- c('DT_confidence', 'DT_confidence', 'DT_RT')
x.var <- c('DT_RT', 'condition*instruction', 'condition*instruction')

for (i in seq_along(y.var)){
  for (j in seq_along(x.var)){
    if (i==j){
      model <- lmer(paste(y.var[i],' ~', x.var[j],' + (1|ID) + (1|item) + (1|trial)'), df)
      print(report(model))
      print(parameters(model, effects = "fixed"))
      if (grepl('\\*', x.var[j]) == TRUE){
        p <- interactions::cat_plot(model, pred = instruction, modx = condition)
      }
      else{
        viz_data <- estimate_relation(model)
        p <- ggplot(data = viz_data, aes(x = DT_RT, y = Predicted)) +
          geom_line() +
          geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3) +
          geom_point(data = df, aes(x = DT_RT, y = DT_confidence, color = ID))
      }
      print(p)
    }
  }
}
```

```{r confidence and RT}
model <- lmer(DT_confidence ~ DT_RT + (1|ID) + (1|item), df) # removed trial
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model)
p <- ggplot(data = viz_data, aes(x = DT_RT, y = Predicted)) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3) +
  geom_point(data = df, aes(x = DT_RT, y = DT_confidence, color = ID))
p
# there's still a singular fit?
# weirdly significant given the plot
# what's the include_random thing (6.5 vs 6.7)
```

```{r condition*instruction}
y.var <- c('DT_confidence', 'DT_RT')

for (i in seq_along(y.var)){
  model <- lmer(paste(y.var[i],' ~ condition*instruction + (1|ID) + (1|item)'), df)
  print(parameters(model, effects = "fixed"))
  means <- modelbased::estimate_means(model, at = c("condition", "instruction"))
  contrasts <- modelbased::estimate_contrasts(model, contrast = c("condition", "instruction"))
  p <- ggplot(means, aes(x = instruction, y = Mean, colour = condition)) + 
    geom_line(aes(group = condition)) +
    geom_pointrange(aes(ymin = CI_low, ymax= CI_high))
  print(p)
}
# what's the better way to loop
# significant interaction for confidence: in truth, polygraph increases confidence but no diff in lie
# non-significant interaction for RT
```

```{r style*instruction}
y.var <- c('DT_confidence', 'DT_RT')

for (i in seq_along(y.var)){
  model <- lmer(paste(y.var[i],' ~ style*instruction + (1|ID) + (1|item)'), df)
  print(parameters(model, effects = "fixed"))
  means <- modelbased::estimate_means(model, at = c("style", "instruction"))
  contrasts <- modelbased::estimate_contrasts(model, contrast = c("style", "instruction"))
  p <- ggplot(means, aes(x = instruction, y = Mean, colour = style)) + 
    geom_line(aes(group = style)) +
    geom_pointrange(aes(ymin = CI_low, ymax= CI_high))
  print(p)
}
# non-significant interaction for both (large CI)
```

```{r}
model <- lmer(DT_confidence ~ condition*instruction + (1|ID) + (1|item), df)
report(model)
parameters(model, effects = "fixed")
means <- modelbased::estimate_means(model, at = c("condition", "instruction"))
contrasts <- modelbased::estimate_contrasts(model, contrast = c("condition", "instruction"))
ggplot(means, aes(x = instruction, y = Mean, colour = condition)) + 
  geom_line(aes(group = condition)) +
  geom_pointrange(aes(ymin = CI_low, ymax= CI_high))
performance::performance(model)
```