### 2021-10-21 3_inferential_questionnaire.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---
## The effect of TOM ability on confidence and RT
### Yoni task
no correlation of yoni score with confidence
```{r}
model <- glmmTMB(DT_confidence ~ yoni_total + DT_RT + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model, at = "yoni_total")
ggplot(data = viz_data, aes(x = yoni_total, y = Predicted)) +
  geom_point(data = df.plot, aes(x = yoni_total, y = DT_confidence, color = ID), show.legend = FALSE) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3)
```

no correlation of yoni score with RT
```{r}
model <- glmmTMB(DT_RT ~ yoni_total + DT_confidence + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model, at = "yoni_total")
ggplot(data = viz_data, aes(x = yoni_total, y = Predicted)) +
  geom_point(data = df.plot, aes(x = yoni_total, y = DT_RT, color = ID), show.legend = FALSE) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3)
```

### BES
no correlation of BES score with confidence
```{r}
model <- glmmTMB(DT_confidence ~ BES_total + DT_RT + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model, at = "BES_total")
ggplot(data = viz_data, aes(x = BES_total, y = Predicted)) +
  geom_point(data = df.plot, aes(x = BES_total, y = DT_confidence, color = ID), show.legend = FALSE) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3)
```

no correlation of BES score with RT
```{r}
model <- glmmTMB(DT_RT ~ BES_total + DT_confidence + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model, at = "BES_total")
ggplot(data = viz_data, aes(x = BES_total, y = Predicted)) +
  geom_point(data = df.plot, aes(x = BES_total, y = DT_RT, color = ID), show.legend = FALSE) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3)
```

## The effect of interoceptive ability on confidence and RT
### HCT
no correlation of HCT accuracy with confidence
```{r}
model <- glmmTMB(DT_confidence ~ HCT_accuracy + DT_RT + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model, at = "HCT_accuracy")
ggplot(data = viz_data, aes(x = HCT_accuracy, y = Predicted)) +
  geom_point(data = df.plot, aes(x = HCT_accuracy, y = DT_confidence, color = ID), show.legend = FALSE) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3)
```

no correlation of HCT accuracy with RT
```{r}
model <- glmmTMB(DT_RT ~ HCT_accuracy + DT_confidence + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model, at = "HCT_accuracy")
ggplot(data = viz_data, aes(x = HCT_accuracy, y = Predicted)) +
  geom_point(data = df.plot, aes(x = HCT_accuracy, y = DT_RT, color = ID), show.legend = FALSE) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3)
```

### MAIA
the higher the MAIA the higher the confidence (p=0.048)
```{r}
model <- glmmTMB(DT_confidence ~ MAIA_total + DT_RT + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model, at = "MAIA_total")
ggplot(data = viz_data, aes(x = MAIA_total, y = Predicted)) +
  geom_point(data = df.plot, aes(x = MAIA_total, y = DT_confidence, color = ID), show.legend = FALSE) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3)
```

the higher the MAIA the lower the RT (p=0.035)
```{r}
model <- glmmTMB(DT_RT ~ MAIA_total + DT_confidence + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model, at = "MAIA_total")
ggplot(data = viz_data, aes(x = MAIA_total, y = Predicted)) +
  geom_point(data = df.plot, aes(x = MAIA_total, y = DT_RT, color = ID), show.legend = FALSE) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3)
```

```{r test}
model <- glmmTMB(DT_confidence ~ MAIA_total + DT_RT + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model, at = "MAIA_total")
ggplot(data = viz_data, aes(x = MAIA_total, y = Predicted)) +
  geom_boxplot(data = df, aes(x = MAIA_total, y = DT_confidence, color = ID), show.legend = FALSE) + 
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3)
```

## idea for within
independent variable: 
each participant is assigned a percentile for yoni/bes/hct/maia
objective-objective and subjective-subjective to obtain difference in TOM-int ability

dependent variable:
average of DT_confidence in social - polygraph
hmm

hypothesis: 
the higher the difference in TOM-int, the higher the social-polygraph confidence
(though this does not seem to be the case in descriptive)