### 2021-11-03 2_inferential_expt.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r}
plot_expt <- function(model, var){
  means <- estimate_means(model, at = c(var, "instruction"))
  ggplot(means, aes_string(x = "instruction", y = "Mean", colour = var)) + 
    geom_line(aes_string(group = var)) +
    geom_pointrange(aes(ymin = CI_low, ymax= CI_high))
}
```

## The effect of confidence on RT
Confidence has a significant impact on RT. Increased confidence decreases RT.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ DT_RT + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
viz_data <- estimate_relation(model)
ggplot(data = viz_data, aes(x = DT_RT, y = Predicted)) +
  geom_point(data = df, aes(x = DT_RT, y = DT_confidence, color = ID), show.legend = FALSE) +
  geom_line() +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.3)
```

## Confidence
### Interaction of condition and instruction
There is a significant interaction of condition and instruction on confidence.
The increase in confidence from lie to truth is more in polygraph than social.
```{r}
model <- glmmTMB(DT_confidence ~ condition*instruction + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_expt(model, "condition")
```

### Interaction of style and instruction
There is no significant interaction of style and instruction on confidence.
```{r}
model <- glmmTMB(DT_confidence ~ style*instruction + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_expt(model, "style")
```

## Reaction time
### Interaction of condition and instruction
There is no significant interaction of condition and instruction on RT.
```{r}
model <- glmmTMB(DT_RT ~ condition*instruction + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_expt(model, "condition")
```

### Interaction of style and instruction
There is no significant interaction of style and instruction on RT.
Indirect style significantly increases RT.
```{r}
model <- glmmTMB(DT_RT ~ style*instruction + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_expt(model, "style")
```