### 2021-11-03 3_inferential_behv.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r}
plot_behv <- function(model, varx, vary) {
  viz_data <- estimate_relation(model, at = c("condition", varx, "instruction"))
  ggplot(data = viz_data, aes_string(x = varx, y = "Predicted")) +
    geom_point(data = df.plot, aes_string(y = vary, color = "condition"), show.legend = FALSE) +
    geom_line(aes(color = condition)) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = condition), alpha = 0.3) + 
    facet_wrap(~instruction)
}
```

## Summary
```{r}
summary <- data.frame("yoni", "decrease (trend)", "social", "increase (ns)", "ns")
names(summary) <- c("measure", "conf main effect (lie)", "conf interact", "RT main effect (lie)", "RT interact")
summary <- summary %>% 
  rbind(c("BES", "decrease (ns)", "social (trend)", "decrease (ns)", "polygraph")) %>% 
  rbind(c("HCT confidence", "increase", "polygraph", "decrease", "ns")) %>% 
  rbind(c("HCT accuracy", "increase", "polygraph", "decrease (ns)", "ns")) %>% 
  rbind(c("HCT awareness", "decrease", "social", "increase", "polygraph")) %>% 
  rbind(c("MAIA", "increase", "ns", "decrease", "social"))
summary
```

## TOM ability
### Yoni task
When instructed to lie, the decrease in confidence in participants with higher yoni score is less in social than polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ instruction/(condition*yoni_total) + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "yoni_total", "DT_confidence")
```

When instructed to lie, participants with higher yoni score show no significant difference in RT between social and polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_RT ~ instruction/(condition*yoni_total) + style + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "yoni_total", "DT_RT")
```

### BES
When instructed to lie, the decrease in confidence in participants with higher BES score has a trend towards being less in social than polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ instruction/(condition*BES_total) + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "BES_total", "DT_confidence")
```

When instructed to lie, the decrease in RT in participants with higher BES score is more in social than polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_RT ~ instruction/(condition*BES_total) + style + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "BES_total", "DT_RT")
```

## Interoceptive ability
### HCT confidence
When instructed to lie, the increase in confidence in participants with higher HCT confidence is more in polygraph than social condition.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ instruction/(condition*HCT_confidence) + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "HCT_confidence", "DT_confidence")
```

When instructed to lie, participants with higher HCT confidence show no significant difference in RT between social and polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_RT ~ instruction/(condition*HCT_confidence) + style + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "HCT_confidence", "DT_RT")
```

### HCT accuracy
When instructed to lie, the increase in confidence in participants with higher HCT accuracy is more in polygraph than social condition.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ instruction/(condition*HCT_accuracy) + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "HCT_accuracy", "DT_confidence")
```

When instructed to lie, participants with higher HCT accuracy show no significant difference in RT between social and polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_RT ~ instruction/(condition*HCT_accuracy) + style + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "HCT_accuracy", "DT_RT")
```

### HCT awareness
When instructed to lie, participants with higher HCT awareness have higher confidence in social than polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ instruction/(condition*HCT_awareness) + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "HCT_awareness", "DT_confidence")
```

When instructed to lie, the increase in RT in participants with higher HCT awareness is less in social than polygraph condition. (Strange considering confidence increases)
```{r warning=FALSE}
model <- glmmTMB(DT_RT ~ instruction/(condition*HCT_awareness) + style + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "HCT_awareness", "DT_RT")
```

### MAIA
When instructed to lie, participants with higher MAIA score show no significant difference in confidence between social and polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ instruction/(condition*MAIA_total) + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "MAIA_total", "DT_confidence")
```

When instructed to lie, the decrease in RT in participants with higher MAIA score is more in polygraph than social condition.
```{r warning=FALSE}
model <- glmmTMB(DT_RT ~ instruction/(condition*MAIA_total) + style + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "MAIA_total", "DT_RT")
```

## Lie scale
### Lie ability
Higher lie ability increases confidence more in lie than truth.
When instructed to lie, participants with higher lie ability show no significant difference in confidence between social and polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ instruction/(condition*lie_ability) + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "lie_ability", "DT_confidence")
```

When instructed to lie, participants with higher lie ability show no significant difference in RT between social and polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_RT ~ instruction/(condition*lie_ability) + style + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "lie_ability", "DT_RT")
```

### Lie frequency
When instructed to lie, participants with higher lie frequency show no significant difference in confidence between social and polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ instruction/(condition*lie_frequency) + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "lie_frequency", "DT_confidence")
```

Participants with higher lie frequency show significant increase in RT for both lie and truth.
When instructed to lie, the increase in RT in participants with higher lie frequency is more in polygraph than social condition.
```{r warning=FALSE}
model <- glmmTMB(DT_RT ~ instruction/(condition*lie_frequency) + style + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "lie_frequency", "DT_RT")
```

### Lie negativity
When instructed to lie, participants with higher perception of lie negativity show no significant difference in confidence between social and polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ instruction/(condition*lie_negativity) + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "lie_negativity", "DT_confidence")
```

Participants with higher perception of lie negativity show significant decrease in RT for both lie and truth.
When instructed to lie, the decrease in RT in participants with higher perception of lie negativity is more in polygraph than social condition.
```{r warning=FALSE}
model <- glmmTMB(DT_RT ~ instruction/(condition*lie_negativity) + style + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "lie_negativity", "DT_RT")
```

### Lie contextuality
When instructed to lie, participants with higher lie contextuality show no significant difference in confidence between social and polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_confidence ~ instruction/(condition*lie_contextuality) + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "lie_contextuality", "DT_confidence")
```

When instructed to lie, the decrease in RT in participants with higher lie contextuality is more in social than polygraph condition.
```{r warning=FALSE}
model <- glmmTMB(DT_RT ~ instruction/(condition*lie_contextuality) + style + (1|ID) + (1|item), data = df)
parameters(model, effects = "fixed")
plot_behv(model, "lie_contextuality", "DT_RT")
```