### 2021-11-11 1_Measures.Rmd
Revise analysis script

---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Theory of Mind / Empathy

### Yoni Task

```{r message=FALSE}
df %>% 
  group_by(Participant) %>% 
  select(starts_with("YONI_")) |> 
  summarise_all(mean, na.rm=TRUE) |> 
  tidyr::pivot_longer(-Participant, values_to = "Scores") |> 
  mutate(name = paste0(str_replace(name, "_", " ("), ")")) |> 
  ggplot(aes(x = Scores, fill = name)) +
  geom_density() +
  scale_fill_manual(values = c("YONI (Affective)" = "Purple",
                                 "YONI (Cognitive)" = "Blue",
                                 "YONI (Physical)" = "Green",
                                 "YONI (Total)"= "DarkBlue"),
                      guide = "none") +
  facet_wrap(~name, scales = "free")
```

### BES Questionnaire

```{r message=FALSE}
df %>% 
  group_by(Participant) %>% 
  select(starts_with("BES_")) |> 
  summarise_all(mean, na.rm=TRUE) |> 
  tidyr::pivot_longer(-Participant, values_to = "Scores") |> 
  mutate(name = paste0(str_replace(name, "_", " ("), ")")) |> 
  ggplot(aes(x = Scores, fill = name)) +
  geom_density() +
  scale_fill_manual(values = c("BES (Affective)" = "Purple",
                               "BES (Cognitive)" = "Blue",
                               "BES (Total)"= "DarkBlue"),
                      guide = "none") +
  facet_wrap(~name, scales = "free")
```

## Interoception

### Heartbeat Counting Task (HCT)

```{r message=FALSE}
df %>% 
  group_by(Participant) %>% 
  select(starts_with("HCT_")) |> 
  summarise_all(mean, na.rm=TRUE) |> 
  tidyr::pivot_longer(-Participant, values_to = "Scores") |> 
  mutate(name = paste0(str_replace(name, "_", " ("), ")")) |> 
  ggplot(aes(x = Scores, fill = name)) +
  geom_density() +
  scale_fill_manual(values = c("HCT (Accuracy)" = "Red",
                               "HCT (Awareness)" = "Orange",
                               "HCT (Confidence)"= "DarkOrange"),
                      guide = "none") +
  facet_wrap(~name, scales = "free")
```

### MAIA

```{r message=FALSE}
df %>% 
  group_by(Participant) %>% 
  select(starts_with("MAIA_")) |> 
  summarise_all(mean, na.rm=TRUE) |> 
  tidyr::pivot_longer(-Participant, values_to = "Scores") |> 
  mutate(name = paste0(str_replace(name, "_", " ("), ")")) |> 
  ggplot(aes(x = Scores, fill = name)) +
  geom_density() +
  scale_fill_brewer(palette = "Reds", guide = "none") +
  facet_wrap(~name, scales = "free")
```

## Deception

### LIE Scale

```{r message=FALSE}
df %>% 
  group_by(Participant) %>% 
  select(starts_with("LIE_")) |> 
  summarise_all(mean, na.rm=TRUE) |> 
  tidyr::pivot_longer(-Participant, values_to = "Scores") |> 
  mutate(name = paste0(str_replace(name, "_", " ("), ")")) |> 
  ggplot(aes(x = Scores, fill = name)) +
  geom_density() +
  scale_fill_manual(values = c("LIE (Ability)" = "#2196F3",
                               "LIE (Frequency)" = "#4CAF50",
                               "LIE (Contextuality)"= "#FF9800",
                               "LIE (Negativity)"= "#E91E63"),
                    guide = "none") +
  facet_wrap(~name, scales = "free")
```

### Deception Task

#### Summary Table

```{r warning=FALSE}

df |> 
  group_by(Participant, Answer) |> 
  summarise(Confidence = paste(insight::format_value(mean(Confidence, na.rm = TRUE)),
                               " +- ",
                               insight::format_value(sd(Confidence, na.rm = TRUE))),
            RT = paste(insight::format_value(mean(RT, na.rm = TRUE)),
                       " +- ",
                       insight::format_value(sd(RT, na.rm = TRUE)))) |> 
  arrange(Participant) |> 
  knitr::kable()
```

#### Outliers

```{r warning=FALSE}
df <- df |> 
  dplyr::filter(Participant != "S9",  # Extreme answers
                !Participant %in% c("S3", "S15", "S19", "S23"))  # No data
```

#### Distributions

```{r warning=FALSE}
p1 <- df |> 
  dplyr::filter(!Participant %in% c("S29")) |>
  ggplot(aes(x = Confidence, fill = Participant)) +
  geom_density(alpha = 0.1) +
  see::scale_fill_material_d(palette = "rainbow", guide = "none") +
  see::theme_modern() +
  scale_x_continuous(labels = scales::percent, expand=expansion(c(0, .05))) +
  scale_y_continuous(expand=expansion(c(0, .05))) +
  facet_wrap(~Answer)
p2 <- df |> 
  dplyr::filter(!Participant %in% c("S29")) |> 
  ggplot(aes(x = RT, fill = Participant)) +
  geom_density(alpha = 0.1) +
  see::scale_fill_material_d(palette = "rainbow", guide = "none") +
  scale_x_continuous(expand=expansion(c(0, .05))) +
  scale_y_continuous(expand=expansion(c(0, .05))) +
  facet_wrap(~Answer)
p1 / p2
```