### 2021-11-11 3_ToM.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Yoni Task {.tabset}

- When instructed to lie, the decrease in confidence in participants with higher yoni score is less in social than polygraph condition.

### Correlation with LIE Scale

```{r message=FALSE, warning=FALSE}
get_correlation <- function(var = "YONI_", var2 = "LIE_") {
  r <- correlation(select(dfsub, starts_with(var2)),
                 select(dfsub, starts_with(var)), 
                 p_adjust = "none") |> 
  mutate(Parameter1 = paste0(str_replace(Parameter1, "_", " ("), ")"),
         Parameter2 = paste0(str_replace(Parameter2, "_", " ("), ")"))

  p <- summary(r) |> 
    plot() +
    theme_minimal()
  list(r = r, plot = p)
}

r <- get_correlation(var = "YONI_")
r$plot
```

### Total Score

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * YONI_Total) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

get_parameters <- function(model) {
  # Parameters
  params <- parameters::parameters(model, effects = "fixed") 
  # Marginal effects
  at <- c("Answer", "Condition")
  trend <- insight::find_predictors(model)$conditional
  trend <- trend[!trend %in% at]
  marg <- modelbased::estimate_slopes(model, trend = trend, at = at)
  # Output
  list(params = params, marginal_effects = marg)
}

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect <- function(model, var = "YONI_Total", outcome = "Confidence") {
  data <- df |> 
    group_by(Participant, Answer, Condition) |> 
    summarise({{var}} := mean(.data[[var]], na.rm = TRUE),
              SD = sd(.data[[outcome]], na.rm = TRUE),
              {{outcome}} := mean(.data[[outcome]], na.rm = TRUE),
              CI_low = .data[[outcome]] - SD / 2,
              CI_high = .data[[outcome]] + SD / 2)

  dodge_width <- 0.02 * diff(range(data[[var]]))
  ylab <- ifelse(outcome == "RT", "Reaction Time (s)", "Confidence")

  link_data <- estimate_relation(model, at = c("Condition", var, "Answer"), length = 30)
  ggplot(link_data, aes_string(x = var, y = "Predicted")) +
    geom_pointrange(data = data, aes_string(y = outcome, color = "Condition", ymin = "CI_low", ymax = "CI_high"), position = position_dodge(width = dodge_width)) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Condition), alpha = 0.33) + 
    geom_line(aes(color = Condition)) +
    labs(y = ylab, x = paste0(stringr::str_replace(var, "_", " ("), ")")) +
    scale_color_manual(values = c("Polygraph" = "#FF5722", "Social" = "#2196F3")) +
    scale_fill_manual(values = c("Polygraph" = "#FF5722", "Social" = "#2196F3")) +
    facet_wrap(~Answer) 
}

plot_effect(model, var = "YONI_Total", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * YONI_Total) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "YONI_Total", outcome = "RT")
```

## BES {.tabset}

### Total Score

#### LIE Scale

```{r message=FALSE, warning=FALSE}
r <- get_correlation(var = "BES_")
r$plot
```

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * BES_Total) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "BES_Total", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * BES_Total) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "BES_Total", outcome = "RT")
```