### 2021-11-11 4_Intero.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Heartbeat Counting {.tabset}

### Correlation with LIE Scale

```{r message=FALSE, warning=FALSE}
r <- get_correlation(var = "HCT_")
r$plot

# plot(correlation::cor_test(dfsub, "HCT_Accuracy", "LIE_Ability"))
```

### Accuracy

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * HCT_Accuracy) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "HCT_Accuracy", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * HCT_Accuracy) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "HCT_Accuracy", outcome = "RT")
```

### Interoceptive Confidence

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * HCT_Confidence) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "HCT_Confidence", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * HCT_Confidence) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "HCT_Confidence", outcome = "RT")
```

### Interoceptive Awareness

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * HCT_Awareness) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "HCT_Awareness", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * HCT_Awareness) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "HCT_Awareness", outcome = "RT")
```

## MAIA {.tabset}

### Correlation with LIE Scale

```{r message=FALSE, warning=FALSE}
r <- get_correlation(var = "MAIA_")
r$plot +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Total Score

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_Total) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "MAIA_Total", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_Total) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "MAIA_Total", outcome = "RT")
```

### Noticing

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_Noticing) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "MAIA_Noticing", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_Noticing) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "MAIA_Noticing", outcome = "RT")
```

### Not Distracting

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_NotDistracting) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "MAIA_NotDistracting", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_NotDistracting) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
plot_effect(model, var = "MAIA_NotDistracting", outcome = "RT")
```