### 2021-11-23 4_Intero.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Heartbeat Counting {.tabset}

### Interoceptive Accuracy

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * HCT_Accuracy) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_hct_accuracy <- plot_effect(model, results, var = "HCT_Accuracy", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * HCT_Accuracy) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_hct_accuracy <- plot_effect(model, results, var = "HCT_Accuracy", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * HCT_Accuracy) + (1|Participant), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_hct_accuracy <- plot_effect(model, results, var = "HCT_Accuracy", outcome = "HeartRate")
```

### Interoceptive Confidence

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * HCT_Confidence) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_hct_confidence <- plot_effect(model, results, var = "HCT_Confidence", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * HCT_Confidence) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_hct_confidence <- plot_effect(model, results, var = "HCT_Confidence", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * HCT_Confidence) + (1|Participant), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_hct_confidence <- plot_effect(model, results, var = "HCT_Confidence", outcome = "HeartRate")
```

### Interoceptive Awareness

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * HCT_Awareness) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_hct_awareness <- plot_effect(model, results, var = "HCT_Awareness", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * HCT_Awareness) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_hct_awareness <- plot_effect(model, results, var = "HCT_Awareness", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * HCT_Awareness) + (1|Participant), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_hct_awareness <- plot_effect(model, results, var = "HCT_Awareness", outcome = "HeartRate")
```

### Correlation with LIE Scale

```{r message=FALSE, warning=FALSE}
r <- get_correlation(var = "HCT_")
r$plot

# plot(correlation::cor_test(dfsub, "HCT_Accuracy", "LIE_Ability"))
```

### Summary

```{r message=FALSE, warning=FALSE, fig.height=figheight*3, fig.width=figwidth*1.5}
p_conf_hct_accuracy / 
  p_conf_hct_confidence / 
  p_conf_hct_awareness +
  patchwork::plot_annotation(title = "Heartbeat Counting Task", theme = theme(plot.title = element_text(hjust = 0.5)))

p_rt_hct_accuracy / 
  p_rt_hct_confidence / 
  p_rt_hct_awareness +
  patchwork::plot_annotation(title = "Heartbeat Counting Task", theme = theme(plot.title = element_text(hjust = 0.5)))

p_hr_hct_accuracy / 
  p_hr_hct_confidence / 
  p_hr_hct_awareness +
  patchwork::plot_annotation(title = "Heartbeat Counting Task", theme = theme(plot.title = element_text(hjust = 0.5)))
```

## Heartbeat Tracking {.tabset}

## MAIA {.tabset}

### Total Score

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_Total) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_maia_total <- plot_effect(model, results, var = "MAIA_Total", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_Total) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_maia_total <- plot_effect(model, results, var = "MAIA_Total", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * MAIA_Total) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_maia_total <- plot_effect(model, results, var = "MAIA_Total", outcome = "HeartRate")
```

### Noticing

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_Noticing) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_maia_noticing <- plot_effect(model, results, var = "MAIA_Noticing", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_Noticing) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_maia_noticing <- plot_effect(model, results, var = "MAIA_Noticing", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * MAIA_Noticing) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_maia_noticing <- plot_effect(model, results, var = "MAIA_Noticing", outcome = "HeartRate")
```

### Not Distracting

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_NotDistracting) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_maia_notdistracting <- plot_effect(model, results, var = "MAIA_NotDistracting", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_NotDistracting) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_maia_notdistracting <- plot_effect(model, results, var = "MAIA_NotDistracting", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * MAIA_NotDistracting) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_maia_notdistracting <- plot_effect(model, results, var = "MAIA_NotDistracting", outcome = "HeartRate")
```

### Body Listening

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_BodyListening) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_maia_bodylistening <- plot_effect(model, results, var = "MAIA_BodyListening", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_BodyListening) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_maia_bodylistening <- plot_effect(model, results, var = "MAIA_BodyListening", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * MAIA_BodyListening) + (1|Participant), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_maia_bodylistening <- plot_effect(model, results, var = "MAIA_BodyListening", outcome = "HeartRate")
```

### Trusting

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_Trusting) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_maia_trusting <- plot_effect(model, results, var = "MAIA_Trusting", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_Trusting) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_maia_trusting <- plot_effect(model, results, var = "MAIA_Trusting", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * MAIA_Trusting) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_maia_trusting <- plot_effect(model, results, var = "MAIA_Trusting", outcome = "HeartRate")
```

### Not Worrying

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_NotWorrying) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_maia_notworrying <- plot_effect(model, results, var = "MAIA_NotWorrying", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_NotWorrying) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_maia_notworrying <- plot_effect(model, results, var = "MAIA_NotWorrying", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * MAIA_NotWorrying) + (1|Participant), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_maia_notworrying <- plot_effect(model, results, var = "MAIA_NotWorrying", outcome = "HeartRate")
```

### Emotional Awareness

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_EmotionalAwareness) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_maia_emotionalawareness <- plot_effect(model, results, var = "MAIA_EmotionalAwareness", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_EmotionalAwareness) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_maia_emotionalawareness <- plot_effect(model, results, var = "MAIA_EmotionalAwareness", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * MAIA_EmotionalAwareness) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_maia_emotionalawareness <- plot_effect(model, results, var = "MAIA_EmotionalAwareness", outcome = "HeartRate")
```

### Attention Regulation

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_AttentionRegulation) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_maia_attentionregulation <- plot_effect(model, results, var = "MAIA_AttentionRegulation", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_AttentionRegulation) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_maia_attentionregulation <- plot_effect(model, results, var = "MAIA_AttentionRegulation", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * MAIA_AttentionRegulation) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_maia_attentionregulation <- plot_effect(model, results, var = "MAIA_AttentionRegulation", outcome = "HeartRate")
```

### Self Regulation

#### Confidence

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(Confidence ~ Answer / (Condition * MAIA_SelfRegulation) + (1|Participant) + (1|Item), 
                 data = df, family = beta_family())

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_conf_maia_selfregulation <- plot_effect(model, results, var = "MAIA_SelfRegulation", outcome = "Confidence")
```

#### RT

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(RT ~ Answer / (Condition * MAIA_SelfRegulation) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_rt_maia_selfregulation <- plot_effect(model, results, var = "MAIA_SelfRegulation", outcome = "RT")
```

#### Heart Rate

```{r message=FALSE, warning=FALSE}
model <- glmmTMB(HeartRate ~ Answer / (Condition * MAIA_SelfRegulation) + (1|Participant) + (1|Item), 
                 data = df)

results <- get_parameters(model)
results$params
results$marginal_effects
```

```{r message=FALSE, warning=FALSE}
p_hr_maia_selfregulation <- plot_effect(model, results, var = "MAIA_SelfRegulation", outcome = "HeartRate")
```

### Correlation with LIE Scale

```{r message=FALSE, warning=FALSE}
r <- get_correlation(var = "MAIA_")
r$plot +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Summary

```{r message=FALSE, warning=FALSE, fig.height=figheight*4, fig.width=figwidth*1.5}
p_conf_maia_total / 
  p_conf_maia_noticing / 
  p_conf_maia_notdistracting / 
  p_conf_maia_bodylistening /
  p_conf_maia_trusting / 
  p_conf_maia_notworrying / 
  p_conf_maia_emotionalawareness / 
  p_conf_maia_attentionregulation / 
  p_conf_maia_selfregulation +
  patchwork::plot_annotation(title = "Interoception (MAIA)", theme = theme(plot.title = element_text(hjust = 0.5)))

p_rt_maia_total / 
  p_rt_maia_noticing / 
  p_rt_maia_notdistracting / 
  p_rt_maia_bodylistening /
  p_rt_maia_trusting / 
  p_rt_maia_notworrying / 
  p_rt_maia_emotionalawareness / 
  p_rt_maia_attentionregulation / 
  p_rt_maia_selfregulation +
  patchwork::plot_annotation(title = "Interoception (MAIA)", theme = theme(plot.title = element_text(hjust = 0.5)))

p_hr_maia_total / 
  p_hr_maia_noticing / 
  p_hr_maia_notdistracting / 
  p_hr_maia_bodylistening /
  p_hr_maia_trusting / 
  p_hr_maia_notworrying / 
  p_hr_maia_emotionalawareness / 
  p_hr_maia_attentionregulation / 
  p_hr_maia_selfregulation +
  patchwork::plot_annotation(title = "Interoception (MAIA)", theme = theme(plot.title = element_text(hjust = 0.5)))
```