### 2021-11-23 index.Rmd

---
title: '**The Place of Interoception and Theory of Mind in Deception**'
subtitle: A Subtitle
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: no
    df_print: default
    code_folding: hide
    code_download: yes
  word_document:
    reference_docx: utils/Template_Word.docx
    highlight: pygments
    toc: no
    toc_depth: 3
    df_print: kable
    number_sections: yes
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '2'
    latex_engine: xelatex
editor_options:
  chunk_output_type: console
bibliography: utils/bibliography.bib
csl: utils/apa.csl
---

<!-- 
!!!! IMPORTANT: run `source("utils/render.R")` to publish instead of clicking on 'Knit'
-->

```{r setup, warning=FALSE, message=TRUE, include=FALSE}
# Set up the environment (or use local alternative `source("utils/config.R")`)
source("https://raw.githubusercontent.com/RealityBending/TemplateResults/main/utils/config.R")
options(dplyr.summarise.inform = FALSE) 
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
theme_set(see::theme_modern())
```

# Introduction

```{r badges, echo=FALSE, message=TRUE, warning=FALSE, results='asis'}
# This chunk is a bit complex so don't worry about it: it's made to add badges to the HTML versions
# NOTE: You have to replace the links accordingly to have working "buttons" on the HTML versions
if (!knitr::is_latex_output() && knitr::is_html_output()) {
  cat("![Build](https://github.com/RealityBending/TemplateResults/workflows/Build/badge.svg)
      [![Website](https://img.shields.io/badge/repo-Readme-2196F3)](https://github.com/RealityBending/TemplateResults)
      [![Website](https://img.shields.io/badge/visit-website-E91E63)](https://realitybending.github.io/TemplateResults/)
      [![Website](https://img.shields.io/badge/download-.docx-FF5722)](https://github.com/RealityBending/TemplateResults/raw/main/word_and_pdf/SupplementaryMaterials.docx)
      [![Website](https://img.shields.io/badge/see-.pdf-FF9800)](https://github.com/RealityBending/TemplateResults/blob/main/word_and_pdf/SupplementaryMaterials.pdf)")
}
```

<!-- Title ideas:-->
<!-- Theory of Body and Others -->

# Packages & Data

## Packages

This document was prepared on `r format(Sys.Date())`.

```{r warning=FALSE, message=TRUE, results='asis'}
library(tidyverse)
library(patchwork)
library(glmmTMB)
library(report)
library(parameters)
library(correlation)
library(modelbased)
library(performance)
library(see)

summary(report::report(sessionInfo()))
```

## Data

```{r warning=FALSE, message=TRUE, results='asis'}
# setwd("C:/Users/user/Desktop/Sputnik/2019-23/DeceptionInteroTom")

df <- read.csv("data/data_combined_wPhysio.csv") %>% 
  mutate(ID = as.factor(paste0("S", ID)),
         condition = as.factor(condition),
         item = as.factor(item),
         style = as.factor(style),
         instruction = as.factor(instruction)) |> 
  #TODO: This renaming should be done at the preprocessing stage
  rename("Participant" = "ID",
         "Condition" = "condition",
         "Item" = "item",
         "Phrasing" = "style",
         "Answer" = "instruction",
         "YONI_Total" = "yoni_total",
         "YONI_Affective" = "yoni_affective",
         "YONI_Cognitive" = "yoni_cognitive",
         "YONI_Physical" = "yoni_physical",
         "BES_Total" = "BES_total",
         "BES_Cognitive" = "BES_cognitive",
         "BES_Affective" = "BES_affective",
         "HCT_Confidence" = "HCT_confidence",
         "HCT_Accuracy" = "HCT_accuracy",
         "HCT_Awareness" = "HCT_awareness",
         "MAIA_Total" = "MAIA_total",
         "MAIA_AttentionRegulation" = "MAIA_attention_regulation",
         "MAIA_BodyListening" = "MAIA_body_listening",
         "MAIA_EmotionalAwareness" = "MAIA_emotional_awareness",
         "MAIA_NotDistracting" = "MAIA_not_distracting",
         "MAIA_NotWorrying" = "MAIA_not_worrying",
         "MAIA_Noticing" = "MAIA_noticing",
         "MAIA_SelfRegulation" = "MAIA_self_regulation",
         "MAIA_Trusting" = "MAIA_trusting",
         "LIE_Ability" = "lie_ability",
         "LIE_Frequency" = "lie_frequency",
         "LIE_Negativity" = "lie_negativity",
         "LIE_Contextuality" = "lie_contextuality",
         "Confidence" = "DT_confidence",
         "RT" = "DT_RT",
         "HeartRate" = "ECG_Rate_MeanRaw_1.5_2.5s") |> 
  mutate(Answer = fct_recode(Answer, Lie = "LIE", Truth = "TRUTH")) |> 
  select(-HCT_guess, -HCT_noguess, -HCT_onebreath) |> 
  dplyr::filter(!Participant %in% c("S3", "S15", "S19", "S23"))  # No data

# Adjustments for beta models
df$Confidence[df$Confidence == 1] <- 0.99999
df$Confidence[df$Confidence == 0] <- 0.00001

cat(paste("The data consists of",
          report::report_participants(df,
                                      participants = "Participant",
                                      sex = "Gender",
                                      age = "Age")))
```

# Measures {.tabset}

```{r child='1_Measures.Rmd'}
```

# Manipulation Checks

```{r child='2_ManipulationChecks.Rmd'}
```

# Theory of Mind / Empathy

```{r child='3_ToM.Rmd'}
```

# Interoception

```{r child='4_Intero.Rmd'}
```

# Full Code

The full script of executive code contained in this document is reproduced here.

```{r full_code, ref.label=knitr::all_labels(), eval=FALSE}
```

# Package References

```{r warning=FALSE, message=FALSE, results='asis'}
report::cite_packages(sessionInfo())
```

# References