### 2022-09-19 paper/manuscript_v0.Rmd
manuscript update/polish

---
title: '**The Role of Interoception and Theory of Mind in Deception**'
shorttitle        : "TITLE"
author:
  - name          : "Dominique Makowski"
    affiliation   : " 1, *"
    address       : "HSS 04-18, 48 Nanyang Avenue, Singapore"
    email         : "dmakowski@ntu.edu.sg"
  - name          : "Zen J. Lau"
    affiliation   : " 1"
  - name          : "Tam Pham"
    affiliation   : " 1"
  - name          : "S.H. Annabel Chen"
    affiliation   : " 1, 2, 3, 4"
    corresponding : yes    # Define only one corresponding author
    address       : "HSS 04-19, 48 Nanyang Avenue, Singapore"
    email         : "annabelchen@ntu.edu.sg"
affiliation:
  - id            : "1"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "2"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "Lee Kong Chian School of Medicine, Nanyang Technological University, Singapore"
  - id            : "4"
    institution   : "National Institute of Education, Nanyang Technological University, Singapore"
abstract: |
output:
  papaja::apa6_word:
    keep_tex: yes
  pdf_document: default
  papaja::apa6_pdf:
    keep_tex: yes
keywords: ""
wordcount: ''
bibliography: ["references.bib"]
floatsintext: yes
figsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
documentclass: apa6
classoption: man
header-includes:
- \usepackage[labelfont=bf, font={color=gray,small}]{caption}
- \usepackage{float}
- \usepackage[document]{ragged2e}
csl: utils/apa.csl
---

\justify

```{r r_setup, include = FALSE, warning=FALSE, message=FALSE}
library("papaja")
library("kableExtra")
options(knitr.kable.NA = 'None')
```

```{=html}
<!-- 
Journals:
Neuroscience Letters https://www.journals.elsevier.com/neuroscience-letters
Biological Psychology https://www.journals.elsevier.com/biological-psychology
Psychophysiology https://onlinelibrary.wiley.com/journal/14698986
Cognitive Neurodynamics https://www.springer.com/journal/11571
Neuroscience & Biobehavioral Reviews https://www.journals.elsevier.com/neuroscience-and-biobehavioral-reviews
-->
```
# Introduction

<!-- Little intro-->

The fabrication and delivery of false information is a sophisticated, and cognitively taxing activity [@abe2007deceiving; @leal2008time ], resulting in physiological arousal [@eskritt2021did] which may require effortful self-regulation to conceal[@krokoszinski2016emotion].

<!-- ToM-->

Much of the current research investigating the process and mechanisms involved in lying has considered capabilities such as Theory of Mind (ToM), the ability to understand and make inferences about other people's mental states and predict their behaviours [@el2017deception; @lee2021lying]. ToM is a critical component of deception, since the liar in question must be able to use their inferences to predict whether the individual accepted as true the false information presented to them [@stewart2019deception]. Children begin using deception during the same period of development in which executive function and ToM develops [@sai2021theory], and children trained in ToM saw an increase in lying behaviour [@ding2015theory]. However, there is a growing link between other cognitive and physiological mechanisms that may play a role in both deception detection, and fabricating successful lies

<!-- Intero in deception research-->

Interoceptive awareness (IA) is the ability to monitor, interpret, integrate, and regulate the sensations and state of one's internal organs, and is closely tied to emotional and cognitive processes [@chen2021emerging; @garfinkel2015knowing]. Greater is IA associated with greater self-regulation [@weiss2014interaction], especially mindfulness practitioners who tend to have greater interoceptive awareness and greater ability to regulate stress and associated physiological indicators [@ardi2021effects; @mcmahon2021mindful].

The growing literature examining the relationship between interoception and deception is rapidly growing area of inquiry. Some research has examined how higher interoceptive accuracy, measured through heart beat detection, has been associated with vasoconstriction when faced with a liar, indicating that IA may aid in discriminating high stakes emotional lies and truths [@gunderson2021body]. It is clear that interoception may indicate on an implicit level that one is being lied to [@ten2019different; @van2014registered], however, there is limited research investigating how interoception may work the other way around, in favour of a liar in regulating their own physiological state that may betray their deception.

Attending to interoceptive cues has been associated with greater willingness to engage in cheating [@williams2016valjean], risk-taking behaviours [@ditto2006visceral; @lenggenhager2013listening]. The findings by @vabba2022interoceptive indicate that high cardiac interoceptive participants were more likely to be dishonest, and to tell higher egoistic lies. There is limited research examining whether individuals with high interoceptive accuracy may be more confident in managing physiological tells associated with emotions like guilt or anxiety that the lie may be discovered through suppression or reappraisal [@fustos2013embodiment; @megias2019emotionally].

<!-- Hypothesis -->

The act of lying is often accompanied by physiological responses such as increased heart rate, increased respiration and sweating. Since fluctuations in key physiological indicators as a means to determine whether the individual is lying is a key principle behind the polygraph test and other lie detection techniques, there is a need for more research examining the relationship between IA, theory of mind, and deception. This research attempts to isolate the effects of theory of mind and interoception, and examine how they interact with deception. Of particular interest is how liars might utilize their interoceptive ability to self regulate in towards a more successful deception, and how this ability may compare or interact with other mechanisms at play such as theory of mind. We hypothesize that participants with greater theory of mind ability will show higher lie confidence, shorter reaction time, and lower physiological arousal in the interrogation condition, while participants in the polygraph condition will demonstrate lower lie confidence, lower reaction time and higher physiological arousal during the polygraph prediction.

# Methods

The plan for this study was preregistered (OSF?). The raw data and analysis script can be found at (GITHUB?).

## Participants

*26* participants (65.4% Females, M~*age*~ = 20.9, *SD* = 2.0) were

recruited VIA HOW

for HUAT

-exclusion criteria

This research was approved by the Institutional Review Board (Reference Number: IRB-) of Nanyang Technological University (NTU).

volunteered for [duration of study] for [compensation provided].

## Measures

### Theory of Mind and Empathy

Since Theory of Mind and empathy are closely related, *The Basic Empathy Scale (BES)* [@jolliffe2006development] was used as a subjective measure for theory of mind. *BES* uses a 40-item Likert scale to rate statements such as, "I can usually work out when people are cheerful" from 1-5. We also implemented the *Yoni Task,* which assesses the ability to judge mental states based on verbal cues, eye gaze and facial expression [@shamay2007dissociable].

### Interoceptive Ability

To measure interoception, participants completed a *Heartbeat Counting Task (HCT)* [@schandry1981heart], where participants count their heart beats without taking their pulse, as well as provide a confidence rating for their estimate. At the same time their true heart rate is recorded and the scores are compared to determine the individual's accuracy. They also completed the *MAIA questionnaire* [@mehling2012multidimensional], where they rated 32 statements such as "I can use my breath to reduce tension" on a 1-5 Likert scale.

Additionally, we implemented the Heartbeat Tracking Task (HTT), in which participants - tap one

### Deception

We used the *LIE Scale*, which is a 44-item questionnaire examining deception as a dispositional trait, and measures the participant propensity towards lie frequency, ability, contextuality and negativity [@makowski2021structure].

#### Deception Task

Participants were randomly allocated(??) between two conditions where they underwent 40 trials of a directed lying task. During the task, participants would be instructed to lie for 20 of the trials to questions such as, "What secondary school did you go to?". Prior to beginning the experiment, researchers collected the answers via questionnaire. In each condition, participants received false feedback through different mediums according to the condition that they were assigned, intended to examine the relationship between theory of mind or interoception and the dependent variables (see fig. X?). Participants were randomized between two conditions which we titled the Interrogation Condition and the Polygraph Condition.

In the Interrogation condition, participants were instructed that following their lie, a live examiner over video feed would appear briefly to judge whether they were truthful or not. In actuality, the video of the examiner shown was pre-recorded and the same for all participants. We hypothesized that participants would will rely on theory of mind abilities to predict whether their lie to the examiner was successfully accepted as true.

The Polygraph Condition followed the same routine, however instead of a receiving feedback from an interrogator, following the lie, a video of physiological signals was shown. We hypothesized when made acutely aware of such signals through false feedback, that participants would rely on interoceptive abilities to be more mindful or intentional about modulating the signals that may give them away. The physiological signals in this condition are also false feedback, and pre-recorded and identical for all participants.

During each trial, the time it took for participants to answer is recorded, and their heart rates are measured using an electrocardiogram. Following the trial, participants gave a confidence rating for how successful they believe their lie or truth was.

During the interrogation condition, participants were exposed to video feedback of an examiner following each question. They were informed that the examiner was live and monitoring their physiological signals remotely to determine the honesty of their answers, however this was false, and the same sequence of videos was show to each participant.

### Demographic Information

The demographic information we collected was age, ethnicity, nationality, dominant hand (for application of EDA electrodes), education level (determined by highest qualification received, or in the midst of achieving), religion and duration lived in Singapore.

## Procedure

Each subject was briefed on the experimental procedure and study aims and signed an informed consent document prior to the study commencing.

questionnaire to obtain truthful answers to directed lying task,

when were the measures like yoni etc taken?

sound attenuated room

Physio recording devices (equipment by huat and who)

Where/how did they provide their answer during directed lying task

Deception task

During the experiment, all participants were connected to an electrocardiogram to monitor their heart would be asked a question, [HOW WAS ANSWER GIVEN?] and then connected for 10 seconds to the condition-unique stimulus. Following the exposure to the stimulus, they would be asked to rate their confidence on a sliding scale before repeating the trial 40 times.

Debriefing was provided following the study.

## Data Analysis

The data analysis was carried out using R.... [CITE STUFF]. The full reproducible analysis script is available [INSERT LINK].

We implimented a linear mixed-effect model where

From the deception task, there were three dependent variables: confidence, reaction time and heart rate. For each dependent variable, we investigated its relationship with theory of mind and interoception abilities as measured by the tests on the previous slide. Participant and the specific question asked were added in as random effects, as they are not independent factors. For example, certain participants may be more confident in lying unrelated to their theory of mind or interoception abilities, or certain questions may have a longer reaction time

To further build on the model, trials for lies and truths were separated using an embedded model and the interaction between the dependent variable and condition was probed. This reveals instances where the effect is greater in one condition, for example a higher total Yoni Task score is correlated with decreased lie confidence, but only in the polygraph condition.

### Data Pre-Processing and Analyses

# Results

INSERT VERY FANCY, VERY PRETTY GRAPHS

Implicit relationship between dependent variables. Inter-Measure Correlation:

-   (confident and reaction time) When lying, the faster they answer, the more confident they are.

-   less confident in lies than truths (expected) Condition interaction, greater effect in polygraph (relying on interoception) likely less intuitive feedback cues?? Slower in polygraph makes sense if less intuitive

-   (HR and RT) no relationship

TOM correlated with trials in polygraph (interoception), not expected(!) because expected it to correlate with Interrogation

higher HCT accuracy and MAIA positively correlated with lie confidence in polygraph condition

No effect for reaction time

BES positively correlated with HR,

HCT accuracy negatively correlated with HR

# Discussion

This study aimed to examine the relationship between Theory of Mind and interoception on lie confidence, reaction time and physiological arousal during deception. This was achieved through a directed lying task in which participants were randomized between two two conditions of false feedback where they were intended to rely on either interoceptive or ToM abilities. Across both conditions, the scores show a larger effect on confidence in the polygraph condition, which was both expected and unexpected. This may be due to the polygraph condition may be harder due to the less intuitive feedback cues. Those better at interoception or perceiving their internal signals would hence be more confident in their lies, as we expected. On the other hand, those with better theory of mind become less confident in the absence of another person from which to obtain feedback cues.

There was no significant relationship between reaction time and heart rate, however lie confidence is negatively correlated with reaction time, which is expected. Generally, participants were less confident in lies than truth, however is effect was greater in the polygraph condition, which is possibly due to less intuitive feedback cues. This effect carries over into reaction time, where participants were also slower to respond in the polygraph condition, which is consistent with less intuitive cues.

To our knowledge, this is the first study that has found a significant correlation between interoceptive awareness and lie confidence. Both interoceptive measures, HCT and MAIA, were strongly correlated. Participants with greater IA, also had greater lie confidence and lower truth confidence in the polygraph condition. The three sub-scales of MAIA which had a significant interaction were *Noticing, Body Listening* and *Attention Regulation*. Since participants were made aware of their cardiac activation through false feedback, we anticipated that those with greater accuracy would demonstrate greater lie confidence. Since research has shown deception is more likely when individuals judge a lower risk of confrontation or being caught [@sip2012if; @sip2013pinocchio], these results may offer support to the literature that implicates IA with immoral behaviour. Thus our research may be in support of the findings from @vabba2022interoceptive, where IA was found to be a significant moderator in participant's propensity toward deception and high egoist lies. Similar findings have also linked increased bodily awareness was linked to a greater likelihood of engaging in immoral behavior such as cheating [@williams2016valjean], and risk-taking behaviours [@ditto2006visceral; @lenggenhager2013listening]. However, more research is required to examine this potential link.

A greater HCT score was associated with decreased lie confidence in the polygraph condition. Since HCT awareness is understood to be the correlation between HCT accuracy and HCT confidence, there are two potential explanations for this. First, a low level of awareness could mean that the participant accurately counted their heartbeats, but were not confidence; or vice versa, where they were inaccurate counting but expressed high confidence. The dual interpretations of this data may indicate its lack of reliability. More research is needed to understand the relationship between lie confidence, interoception and potential implications on lying behaviour.

HCT accuracy is negatively correlated with heart rate, but only in the interrogation condition. While it is expected that those with greater interoceptive ability would be better at perceiving and regulating physiological arousal [@megias2019emotionally], we expected to see this result in the polygraph condition which emphasizes interoceptive abilities. A potential explanation for this is that the presence of another person, whom they believe is observing them is understood to increase an individual's interoceptive accuracy [@hazem2017know; @isomura2020direct], which may in turn, cause the participant to attempt to regulate their cardiac response.

Greater ToM abilities negatively correlated with lie confidence, and saw an increase in truth confidence, but only in the polygraph condition. In the Yoni Task, this effect is driven by the cognitive and physical domains. BES saw the effect driven by both the cognitive and affective domain for lie confidence, meanwhile, truth confidence is only driven by the affective domain.

We anticipated that ToM measures would negatively correlate with physiological arousal, however while the Yoni Task had no association with heart rate, we found that BES scores were positively correlated with heart rate for both lies and truths, with a larger effect in the interrogation condition. Although initially unexpected, this finding is makes sense when interpreted in line with past research which finds a close positive correlation between empathy and increased cardiac response [@jauniaux2020emotion; @lischke2018inter]. The lack of correlation between heart rate and the Yoni task, indicates that the elevated cardiac response may be due exclusively to empathy, and attributable to ToM.

## Limitations and Future Research

Potential limitations worth noting in this research is that the interrogation and polygraph conditions may not bring out Theory of Mind and interoceptive abilities to the greatest degree. This must be considered as a potential limitation since the participants were still connected to the ECG in the interrogation condition, which may amplify their interoceptive awareness beyond what they may normally experience during a real-world act of deception. Furthermore, it is possible that there could be an interplay between Theory of Mind and interoception. More research is needed to investigate this relationship in the context of deception.

# Funding

# Author Contributions

# Conflict of Interest Statement

# Acknowledgements

\newpage

# References

```{r create_r-references}
r_refs(file = "references.bib")
```

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs}
:::

```{=tex}
\endgroup
```