### 2022-10-04 4_Intero.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Heartbeat Counting {.tabset}

- Confidence: effect in all.
- RT: effect in confidence only.
- Heart rate: effect in accuracy and awareness.

### Confidence

```{r message=FALSE, warning=FALSE}
results <- data.frame()
for (var in c("HCT_Accuracy", "HCT_Confidence", "HCT_Awareness")) {
  model <- glmmTMB(as.formula(paste0(
    "Confidence ~ Answer / (Condition / ",
    var,
    ") + (1|Participant) + (1|Item)"
  )),
  data = df,
  family = beta_family()
  )

  results <- parameters::parameters(model, effects = "fixed", keep = var) |>
    mutate(Parameter = str_remove(Parameter, "Answer|Condition")) |>
    separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
    data_relocate(select = "Variable", before = 1) |>
    rbind(results)
}

display(results, stars = TRUE)
```

### RT

```{r message=FALSE, warning=FALSE}
results <- data.frame()
for (var in c("HCT_Accuracy", "HCT_Confidence", "HCT_Awareness")) {
  model <- glmmTMB(as.formula(paste0(
    "RT ~ Answer / (Condition / ",
    var,
    ") + (1|Participant) + (1|Item)"
  )),
  data = df
  )

  results <- parameters::parameters(model, effects = "fixed", keep = var) |>
    mutate(Parameter = str_remove(Parameter, "Answer|Condition")) |>
    separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
    data_relocate(select = "Variable", before = 1) |>
    rbind(results)
}

display(results, stars = TRUE)
```

### Heart Rate

```{r message=FALSE, warning=FALSE}
results <- data.frame()
for (var in c("HCT_Accuracy", "HCT_Confidence", "HCT_Awareness")) {
  model <- glmmTMB(as.formula(paste0(
    "HeartRate ~ Answer / (Condition / ",
    var,
    ") + (1|Participant) + (1|Item)"
  )),
  data = df
  )

  results <- parameters::parameters(model, effects = "fixed", keep = var) |>
    mutate(Parameter = str_remove(Parameter, "Answer|Condition")) |>
    separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
    data_relocate(select = "Variable", before = 1) |>
    rbind(results)
}

display(results, stars = TRUE)
```

### Correlation with LIE Scale

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  select(
    Participant,
    starts_with("LIE_"),
    starts_with("HCT_")
  ) |>
  group_by(Participant) |>
  summarise_all(mean)

correlation(select(dfsub, starts_with("LIE_")), select(dfsub, starts_with("HCT_")), p_adjust = "none")
```

## MAIA {.tabset}

- Confidence: effect in total, noticing, not distracting, body listening, not worrying, attention regulation and self regulation.
- RT: effect in total, noticing, body listening, attention regulation and self regulation.
- Heart rate: effect in trusting and emotional awareness only.

### Confidence

```{r message=FALSE, warning=FALSE}
results <- data.frame()
for (var in c(
  "MAIA_Noticing", "MAIA_NotDistracting", "MAIA_NotWorrying",
  "MAIA_AttentionRegulation", "MAIA_EmotionalAwareness", "MAIA_SelfRegulation",
  "MAIA_BodyListening", "MAIA_Trusting", "MAIA_Total"
)) {
  model <- glmmTMB(as.formula(paste0(
    "Confidence ~ Answer / (Condition / ",
    var,
    ") + (1|Participant) + (1|Item)"
  )),
  data = df,
  family = beta_family()
  )

  results <- parameters::parameters(model, effects = "fixed", keep = var) |>
    mutate(Parameter = str_remove(Parameter, "Answer|Condition")) |>
    separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
    data_relocate(select = "Variable", before = 1) |>
    rbind(results)
}

display(results, stars = TRUE)
```

### RT

```{r message=FALSE, warning=FALSE}
results <- data.frame()
for (var in c(
  "MAIA_Noticing", "MAIA_NotDistracting", "MAIA_NotWorrying",
  "MAIA_AttentionRegulation", "MAIA_EmotionalAwareness", "MAIA_SelfRegulation",
  "MAIA_BodyListening", "MAIA_Trusting", "MAIA_Total"
)) {
  model <- glmmTMB(as.formula(paste0(
    "RT ~ Answer / (Condition / ",
    var,
    ") + (1|Participant) + (1|Item)"
  )),
  data = df
  )

  results <- parameters::parameters(model, effects = "fixed", keep = var) |>
    mutate(Parameter = str_remove(Parameter, "Answer|Condition")) |>
    separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
    data_relocate(select = "Variable", before = 1) |>
    rbind(results)
}

display(results, stars = TRUE)
```

### Heart Rate

```{r message=FALSE, warning=FALSE}
results <- data.frame()
for (var in c(
  "MAIA_Noticing", "MAIA_NotDistracting", "MAIA_NotWorrying",
  "MAIA_AttentionRegulation", "MAIA_EmotionalAwareness", "MAIA_SelfRegulation",
  "MAIA_BodyListening", "MAIA_Trusting", "MAIA_Total"
)) {
  model <- glmmTMB(as.formula(paste0(
    "HeartRate ~ Answer / (Condition / ",
    var,
    ") + (1|Participant) + (1|Item)"
  )),
  data = df
  )

  results <- parameters::parameters(model, effects = "fixed", keep = var) |>
    mutate(Parameter = str_remove(Parameter, "Answer|Condition")) |>
    separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
    data_relocate(select = "Variable", before = 1) |>
    rbind(results)
}

display(results, stars = TRUE)
```

### Correlation with LIE Scale

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  select(
    Participant,
    starts_with("LIE_"),
    starts_with("MAIA_")
  ) |>
  group_by(Participant) |>
  summarise_all(mean)

correlation(select(dfsub, starts_with("LIE_")), select(dfsub, starts_with("MAIA_")), p_adjust = "none")
```