### 2023-06-22 paper/manuscript.Rmd
R1 update

---
title: '**The Heart can Lie: A Preliminary Analysis of the Role of Interoception and Theory of Mind in Deception**'
shorttitle        : "Deception, Interoception, and ToM"
author:

  - name          : "Dominique Makowski"
    affiliation   : "1, 2"
    corresponding : yes    # Define only one corresponding author
    address       : "Pevensey 1, University of Sussex, Brighton, UK"
    email         : "D.Makowski@sussex.ac.uk"
    orcid         : 0000-0001-5375-9967
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Data curation"
      - "Formal Analysis"
      - "Funding acquisition"
      - "Investigation"
      - "Methodology"
      - "Project administration"
      - "Resources"
      - "Software"
      - "Supervision"
      - "Validation"
      - "Visualization"
      - "Writing – original draft"
  - name          : "Zen J. Lau"  # lauzenjuen@gmail.com
    affiliation   : " 2"
    role:
      - "Data curation"
      - "Project administration"
      - "Resources"
      - "Investigation"
  - name          : "Tam Pham"  # phamttam17@gmail.com
    affiliation   : " 2"
    role:
      - "Data curation"
      - "Project administration"
      - "Resources"
      - "Investigation"
  - name          : "An Shu Te" # anshu.te@ntu.edu.sg
    affiliation   : "2"
    orcid         : 0000-0001-8008-2824
    role:
      - "Project administration"
      - "Resources"
      - "Investigation"
      - "Writing – original draft"
  - name          : "Stephanie Kirk"  # stephanie.morgank@gmail.com
    affiliation   : "2"
    orcid         : 0000-0002-9312-5552
    role:
      - "Project administration"
      - "Resources"
      - "Writing – original draft"
  - name          : "Claudia Liauw"  # YLIAUW001@e.ntu.edu.sg
    affiliation   : "2"
    role:
      - "Data curation"
      - "Formal Analysis"
  - name          : "S.H. Annabel Chen"  # AnnabelChen@ntu.edu.sg
    affiliation   : "2, 3, 4, 5"
    orcid         : 0000-0002-1540-5516
    role:
      - "Project administration"
      - "Supervision"
      - "Writing – review & editing"

affiliation:
  - id            : "1"
    institution   : "School of Psychology, University of Sussex, UK"
  - id            : "2"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
  - id            : "4"
    institution   : "Lee Kong Chian School of Medicine, Nanyang Technological University, Singapore"
  - id            : "5"
    institution   : "National Institute of Education, Nanyang Technological University, Singapore"
abstract: |
  While a large part of the deception literature focuses on lying detection, the factors contributing to one's ability to lie remain unclear. The present study (n=26) examined the contribution of Theory of Mind (ToM) and interoception on our ability to lie using a directed lie paradigm with two conditions ("Interrogation" and "Polygraph"), designed to enhance each of the two mechanisms. In view of the relatively small sample size, we performed a feature reduction on the measures used to assess our two groups of predictor variables using factor analysis and extracted a unique factor of ToM as well as 4 components of interoception (*Metacognition*, *Listening*, *Focus* and *Regulation*). Using a mixed-design Bayesian analysis, we show that various facets of interoceptive abilities are positively related to the self-rated confidence in one's own lies, especially when under the belief that bodily signals are being monitored (i.e., in the "Polygraph" condition). Beyond providing evidence for the role of the body in lying and raising interesting questions for deception science, these results carry practical implications for criminology and lie detection protocols.
output:
  papaja::apa6_word:
    keep_tex: yes
keywords: "Deception; Interoception; Theory of Mind; Poygraph; Lying Ability"
wordcount: '5004'
bibliography: ["references.bib"]
floatsintext: yes
figsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: yes
draft: no
documentclass: apa6
classoption: man
header-includes:
- \usepackage[labelfont=bf, font={color=gray,small}]{caption}
- \usepackage{float}
- \usepackage[document]{ragged2e}
csl: utils/apa.csl
---

\justify

```{r r_setup, include = FALSE, warning=FALSE, message=FALSE}
library("papaja")
library("kableExtra")
options(knitr.kable.NA = 'None')
```

<!-- 
Journals:
Neuroscience Letters https://www.journals.elsevier.com/neuroscience-letters
Biological Psychology https://www.journals.elsevier.com/biological-psychology
Psychophysiology https://onlinelibrary.wiley.com/journal/14698986
Cognitive Neurodynamics https://www.springer.com/journal/11571
Neuroscience & Biobehavioral Reviews https://www.journals.elsevier.com/neuroscience-and-biobehavioral-reviews
-->

<!-- Little intro-->

Lying - the intentional attempt at instilling a false belief in others [@sip2012if] - is a prevalent phenomenon carrying potentially important consequences. Interestingly, evidence suggests that the successful detection of a lying attempt depends more on the ability of the liar, than on the performance of the lie detector [@verigin2019lie; @bond2008individual; @levine2011sender]. However, with most of the deception literature focused on deception detection [@viji2022history; @sternglanz2019review; @masip2017deception], the factors contributing to one's ability to lie remain unclear. Nevertheless, some findings suggest a relationship between the propensity to tell lies, and traits that characterize the socially malevolent profile known as the Dark Triad [@paulhus2002dark], such as narcissism [@zvi2018correlates] and psychopathy [@rassin2023psychopathy]. While often conceptualized to be immoral and unconscionable, lying is ubiquitous in everyday life, and being able to lie skillfully can sometimes facilitate interpersonal relationships, helping us avoid conflict or causing emotional harm to others [@levine2022prosocial]. In fact, recent research shows that certain forms of deception, such as prosocial lies (i.e., false statements told to benefit others; @levine2022prosocial), can increase trust [@levine2015prosocial]. Moreover, individuals who told altruistic lies were perceived as more benevolent than those who were honest [@levine2014liars].

<!-- research shows that people tend to under-estimate their ability to successfully lie [@elaad2003effects] -->

<!-- ToM  -->

As deception requires the liar to intentionally manipulate the beliefs of others [@burgoon1994interpersonal; @sip2012if], a significant line of research has been focused on the role of theory of mind (ToM) in lying ability. ToM refers to the ability to infer that others have mental states, such as beliefs, emotions and intentions, distinct from ourselves [@lee2021lying; @wellman2001meta; @baron1997mindblindness]. The ability to tell lies, as well as their complexity, have previously been found to be related to higher ToM abilities [@talwar2007lying; @evans2011verbal; @talwar2017role]. However, studies investigating the link between ToM and deception have predominantly been focused on children and neuroatypical individuals [@beaudoin2020systematic; @bora2017meta; @roheger2022progression], and its importance in healthy adults remains to be clarified.

<!-- Intero in Deception research -->

Besides paying attention to the person we lie to, gauging whether they believe us, some attention is also directed inwards: monitoring our own body and its reactions (e.g., cardiac activity and its related changes such as blushing), which could be used as cues to infer our real intent. This begs the question of the potential role of interoceptive abilities in deception ability. Broadly defined as one's sensitivity to their own internal signals and bodily states [@murphy2019classifying; @chen2021emerging; @weiss2014interaction], @garfinkel2015knowing conceptualize interoception as a three-dimensional construct comprising three distinct facets, namely, interoceptive accuracy - the objective ability to monitor internal bodily signals; interoceptive sensibility - the subjective confidence in one's interoceptive accuracy; and interoceptive awareness (also referred to as interoceptive insight; @khalsa2018interoception) - one's ability to correctly evaluate their interoceptive ability. Interoception has increasingly been tied to subjective perceptual experiences [@connell2018interoception; @seth2012interoceptive], as well as individual differences in executive functions, emotional processing, and decision-making [@barrett2015interoceptive; @petzschner2021computational; @murphy2019classifying]. In particular, neuroscientific findings investigating the correlates of interoception have underlined the potential role of the anterior cingulate cortex (ACC) and anterior insula (AI) [@wang2019anterior; @craig2009you; @critchley2004neural; @khalsa2009pathways], both of which are often thought to be activated during deception [@abe2011brain; @sip2008detecting; @baumgartner2013honest], and have been implicated in cognitive processes associated with deception (e.g., cognitive control [@molnar2022anterior], conflict detection [@kerns2004anterior]).

Although few studies exist that investigate the relationship between interoception and deceptive ability *per se*, previous decision-making studies have demonstrated a negative correlation between interoceptive awareness, a metacognitive dimension of interoception, and one's likelihood to make risky decisions [@dunn2010listening; @furman2013interoceptive]. This is in line with the somatic marker hypothesis, which posits that an accurate evaluation of one's bodily signals facilitates the use of such interoceptive feedback to guide rational decision making [@damasio1996somatic]. Indeed, @sugawara2020effects further reports that individuals who received interoceptive training were more likely to show higher interoceptive accuracy and make reasoned decisions. Given that deciding to lie generally involves a consideration of the potential costs of getting caught, and hence could also be perceived as risky behavior [@kireev2013possible], interoception could be construed to be negatively related to lying ability. However, some studies have instead found heightened interoceptive attention (one's self-focus towards internal bodily signals), to predict apathetic, immoral behaviour, such as cheating [@ditto2006visceral; @williams2016valjean; @lenggenhager2013listening]. Extending these findings to social cognition, @vabba2022interoceptive further reports individuals with lower interoception told significantly less egoistic lies when the social reputational stakes were high, whereas individuals with higher interoception did not exhibit a significant difference in the number of lies told. Given the scarce research on interoception and deception, more studies are herein needed to clarify these mixed findings.

<!-- Hypothesis -->

The aim of the present study was to explore the contribution of ToM and interoception abilities on individuals' deception skills, as indicated by their lying confidence, physiological arousal and response time. To this end, we designed a directed-lying paradigm with 2 conditions differing in the nature of their feedback cues. The *Interrogation* condition was designed to emphasize (and preferentially mobilize) ToM-related mechanisms, whereas the *Polygraph* condition was designed to emphasize interoceptive mechanisms. In particular, we expected lying ability (i.e., higher lie confidence, shorter response time and lower physiological arousal), to be positively predicted by individuals' interoceptive abilities in the *Polygraph* condition, and by ToM skills in the *Interrogation* condition. Consistent with the cognitive load approach outlined in several theories of deception (such as the Four-Factor Theory [@riggio1987verbal] and Activation-Decision-Construction Model [@walczyk2014social]), as well as previous findings which suggest response time as a reliable cue to deception [@walczyk2009cognitive; @gonzalez2019can], we regarded shorter response times as a proxy of better lying ability.

# Methods

## Participants

30 university students from Singapore were recruited through posters, flyers, and online social media platforms. 
<!-- Inclusion criteria included residing in Singapore, and an absence of current or history of neurological, psychiatric, or cardiovascular disorders.  -->
Four participants were excluded as their data was not recorded due to technical issues. The final sample consists of 26 participants (Mean age = 20.9, SD = 2.0, range:[18, 25], Sex: 65.4% women, 34.6% men). The heart rate of one participant and response time of one participant were excluded from further analysis due to extreme outlying values. To maximize statistical power, the problematic data from these 2 participants were only excluded from analyses involving those measures; all other data were retained for analyses.

This study was approved by the NTU Institutional Review Board (NTU-IRB-2020-09-007). All participants provided their informed consent prior to participation and were awarded with academic credits upon completion of the study.

## Measures

### Theory of Mind (ToM)

Two measures of ToM and its related constructs were administered. The Yoni Task [@shamay2007dissociable] is a behavioral task which assesses first and second-order ToM abilities in both cognitive and affective domains. Participants were presented with the face of a character named "Yoni", surrounded by 4 colored pictures of objects or faces, one in each corner of the screen. In total, each participant completed 101 trials - 49 trials assessing their affective ToM abilities, 37 trials assessing their cognitive ToM abilities and 15 control trials. During each trial, participants were shown a question pertaining the item Yoni is referring to, and asked to make responses based on specific corresponding cues such as the directions of Yoni's eye gaze, facial expressions etc., In the control trials, participants made judgements based on Yoni's physical context (physical ToM). More specifically, in first-order trials, participants were instructed to make inferences about Yoni's mental state with regards to the objects surrounding it (e.g., "Yoni is thinking of..."). In more complex second-order trials, participants had to correctly infer the interaction between Yoni and others' mental states (e.g., Yoni is thinking of the fruit that ... wants").

The Basic Empathy Scale [BES, @jolliffe2006development], a 20-item self-report questionnaire measuring two dimensions of empathy (cognitive and affective) using a 5-point Likert scale was administered. Although ToM and empathy are regarded as distinct psychological constructs, previous research findings point to them being closely related [@gallant2020developing; @sebastian2012neural]. Specifically, empathy is often thought to be an integral component in the affective dimension of ToM (i.e., the ability to infer what someone else is feeling) [@shamay2010role].
<!-- Interestingly, some studies suggest that people's performance on interpersonal ToM tasks can be disassociated from that of interpersonal ToM [@dvash2014theory; @tine2012unique].  -->

### Interoception

To assess participants' interoceptive ability, participants completed a Heartbeat Counting Task [HCT, @schandry1981heart] while having their actual heartbeats recorded. During the HCT task, participants were instructed to count the number of heartbeats over 5 trials with varying time intervals (20s, 25s, 30s, 35s, 40s), the  order of which was randomized. Interoceptive accuracy was computed from the difference between the estimated number and the real number of heart beats. Interoceptive sensibility was estimated as the average of the confidence ratings presented at the end of each trial. Interoceptive awareness was indexed by the correlation between the objective accuracy and the subjective confidence.

Given its multidimensional nature, the MAIA-2 [@mehling2012multidimensional], a 37-item questionnaire which measures 8 distinct facets of interoception including Noticing (**I notice when I am uncomfortable in my body**), Not-Distracting (**I try to ignore pain**), Not-Worrying (**I can stay calm and not worry when I have feelings of discomfort or pain**), Attention Regulation (**I can refocus my attention from thinking to sensing my body**), Emotional Awareness (**I notice how my body changes when I am angry**), Self-Regulation (**I can use my breath to reduce tension**), Body Listening (**I listen to information from my body about my emotional state**), and Trust (**I trust my body sensations**), using 5-point Likert scales, was also administered.

### Deception

<!-- **TODO: create a procedure figure** -->

Using PsychoPy [@peirce2019psychopy2], we implemented a directed-lying task in which participants were instructed to answer 80 questions pertaining to their personal preferences and subjective experiences, by either lying or telling the truth (depending on whether they see "lie" or "truth" written on the screen). Importantly, they were told that for half of the trials, they would have to convince another participant that would be observing them from a separate room via a webcam connection (*Interrogation* condition). For the remaining trials, participants were tasked to convince a "lie detection machine" that would be assessing their behaviour through their physiological signals (*Polygraph* condition). The two conditions were presented in a counter-balanced order, and each comprised of 40 trials (20 truth; 20 lies).

The sequence of each trial was the same for both conditions. Participants were first instructed to respond to a question shown on a computer screen by verbally lying or telling the truth (for half of the trials in each condition, i.e., n = 20). In addition, as past studies have found associations between lying behaviour and type of question phrasing [@walczyk2022err], each question was phrased either directly (e.g., "What is your favourite sport?"), or indirectly (e.g., "Is your favourite sport Hockey?") to reduce possible confounding effects. Following a short interval (0.7 - 1.5s) to allow time for response preparation, during which "Connecting..." was shown on the screen, a social or bio-feedback cue (for *Interrogation* and *Polygraph* conditions respectively) was displayed for a maximum of 10s or until a response was given. Specifically, participants had to provide their answers verbally, and pressed the space key to signify the end of their response. After another short interval (1.5-2.5s), during which "Disconnecting..." was presented on the screen, participants were asked to rate their confidence in how convincing (i.e., likely to get judged as truthful) they perceived their response to be on a visual analog scale. 
<!-- individuals experience heightened physiological arousal when the items are embedded with concealed information known to them (i.e., indirect questions) regardless of the truthfulness of their response [@verschuere2006psychopathy] -->

In the *Interrogation* condition, participants had to provide their answer while receiving social feedback in the form of a video stream of an examiner, and were informed that the examiner would be evaluating the truthfulness of their responses after observing them through the live video feed. In actuality, the displayed video feed were pre-recorded video clips of a confederate's face (staying still with minimal reactions, and with a medical mask, as the experiment was run during COVID restrictions), and the same video-clips were used with all participants. All participants reported believing that the stream was real and that the examiner was really there during the debriefing.

In the *Polygraph* condition, participants had to provide their answer while receiving bio-feedback in the form of physiological signals (ECG, EDA and Respiration), of which they were informed was live feedback of their own physiological activity. In actuality, we displayed pre-recorded video clips of a confederate's physiological signals. All participants reported believing that their own signals were shown during the debriefing.

Finally, on top of the deception task, we also measured participants' self-reported tendency to lie in their everyday life using the *Lie Scale* [@makowski2021structure], a 16-item questionnaire that assesses 4 dispositional lying dimensions (Ability, Negativity, Contextuality and Frequency).

<!-- ### Demographic Information -->

<!-- We further collected demographic information including age, ethnicity, nationality, handedness (for application of EDA electrodes), religion, duration lived in Singapore as well as highest education qualification achieved (or in the midst of achieving).  -->

## Procedure

A within-subjects design was used in the present study, which is comprised of 2 sessions, to investigate the roles interoception and ToM play in lying ability. During session 1, participants answered a brief demographic survey as well as a questionnaire regarding their personal preferences and subjective experiences (Autobiographical Memory Questionnaire, AMQ), followed by a series of psychological scales (i.e., BES, MAIA and Lie scale), which were randomly displayed. During session 2, cognitive-behavioural tasks (i.e., the deception task, HCT and the Yoni task) were administered to participants while their physiological signals (including cardiac activity (ECG), respiration (RSP) and electrodermal activity (EDA)) were being recorded. The physiological recording devices were set up as follows: ECG was recorded with 3 electrodes placed according to a modified Lead II configuration [@takuma1995alternative], and respiration was measured using a respiration belt. All signals were recorded at 1000Hz via the BioPac MP160 system (BioPac Systems Inc., USA). For all participants, session 2 began with the deception task, followed by the Yoni task and the HCT, with the latter two presented in a randomized order. In the directed-lying task, items in the AMQ were presented as stimuli, with participants' recorded responses (in session 1) used to establish the ground truth. Three outcome variables were recorded for each trial of the deception task, namely the participants' confidence ratings that their answers (lies *vs.* truths) were convincing, the response time (RT) between the question onset and the participant's key press (indicating the end of their verbal answer), and the change in heart rate associated with the response (within a window of 3.5 s).

## Data Analysis

Aware of the low number of participants, we tried to take every step to 1) maximize power by using all available data (from individual trials) with appropriate statistical tools and 2) ensure the robustness of results by cross-validating the findings across different measures and approaches.

Firstly, a manipulation check was carried out to ensure that our outcome variables were sensitive to the experimental manipulations, by testing the effect of the question phrasing (direct *vs.* indirect) and condition (polygraph *vs.* interrogation) on the outcome variables. This analysis was performed using mixed models with the participants and questions both entered as random factors. Marginal contrasts analysis (denoted by $\Delta$) was also performed to clarify the differences between conditions. To allow for a better quantification of the uncertainty associated with the effects, as well as to increase the robustness to outliers and artefactual findings, all statistics were undertaken under the Bayesian framework [@makowski2019bayestestr], using informative priors centred around 0 ($t_{Confidence}(1, 0, 1)$, $t_{RT}(1, 0, 3)$, $t_{Heartrate}(1, 0, 8)$).

To maximize the signal-to-noise ratio, we performed a feature reduction on our two groups of predictor variables (namely, ToM and interoception) using factor analysis over PCA, as the goal was to extract meaningful and consistent factors, rather than merely maximizing the variance explained. Then, we modelled the relationship between these inter-individual composite scores (note that the analysis for all individual variables is nonetheless included in the analysis report) and the 3 outcome variables in interaction with the condition (polygraph *vs.* interrogation). Finally, we investigated the relationship between the deception scale traits, and the ToM and interoception scores using Bayesian correlations. As all the analyses and data have been made available, we will in the manuscript focus on significant, i.e., - in this context - statistically reliable and in our opinion theoretically relevant.

The data analysis was carried out using *R 4.2* [@RCoreTeam2022], *brms* [@Burkner2017], and the *easystats* collection of packages [@makowski2019bayestestr; @correlationArticle; @performanceArticle; @insightArticle], and the physiological signal processing was done using the default routines avaialble in *NeuroKit2* [@makowski2021neurokit2]. Note that EDA was not further analyzed as most participants did not yield any skin conductance responses - which we believe was partly caused by the low temperature (with dry air-con air) of the experimental room.

The analysis was not pre-registered (stemming out from an undergraduate's final year project), but the full reproducible analysis script, statistical results report, and data, are available at 
**[masked for blinding]**
<!-- **https://github.com/DominiqueMakowski/DeceptionInteroTom** -->

# Results

## Manipulation Check

Compared to truth, lies were rated with less confidence ($\Delta=-1.35,~95\%~CI~[-1.46, -1.23],~pd = 100\%$), but no significant difference between the conditions was found. On the other hand, the RT did not differ between truth and lies, but was significantly slower in the polygraph condition for both conditions ($\Delta=0.25,~95\%~CI~[0.62, 0.41],~pd = 100\%$). The heart rate was significantly more elevated during lies as compared to truth ($\Delta=1.16,~95\%~CI~[0.57, 1.73],~pd = 100\%$), and during interrogation as compared to the polygraph condition ($\Delta=4.84 ,~95\%~CI~[4.23, 5.44],~pd = 100\%$).

The indirect phrasing of the question only had a significant effect on RT ($\beta=0.36,~95\%~CI~[0.21, 0.51],~pd = 100\%$), leading to slower answers, regardless of whether they were lies or truths.

## Feature Reduction

The 3 Yoni-task dimensions and the 2 BES traits were combined into a unique factor, labelled *ToM* (explaining 35.76% of variance). It was loaded by the cognitive (.89), affective (.77), physical (.45) Yoni dimensions, and the affective (.41) and cognitive (.17) facets of the BES.

The 8 MAIA dimensions and the 3 HCT components were reduced to 4 factors (explaining 65.17% of variance). The first factor, labelled *Interoception - Meta* (23.59%), was loaded primarily by Attention Regulation (.97), Self-regulation (.63), Emotional awareness (.60), and Noticing (.49) dimensions of the MAIA and the HCT confidence score (.40). The second factor, labelled *Interoception - Listening* (18.54%), was primarily loaded by the Body Listening (.92) and Trusting (.53) MAIA dimensions, and the Awareness (-.60) and Confidence (.46) HCT scores. The third factor, labelled *Interoception - Focus* (12.07%), was primarily loaded by MAIA Not-Distracting (.87), Emotional Awareness (-.40) and HCT Accuracy (.33). The fourth factor, labelled *Interoception - Regulation* (10.97%), was primarily loaded by MAIA not-worrying (.71), HCT Accuracy (.61) and MAIA Trusting (.40). 
<!-- However, due to the interpretation difficulty, we will refer to these dimensions with their main correlates in **Figure 1** [**Note to fix caption for Figure 1 from HCT Trusting to MAIA trusting**]. -->

## Theory of Mind

The composite *ToM* score was significantly associated with less confident ($\beta=-0.19,~95\%~CI~[-0.36, -0.02],~pd = 98.47\%$) and slower lies ($\beta=0.42,~95\%~CI~[0.01, 0.83],~pd = 97.67\%$), specifically in the polygraph condition. No significant effect was found with regards to heart rate, and no correlation was observed with dispositional lying traits (**Figure 1**).

```{r message=FALSE, warning=FALSE, out.width="100%", fig.cap="Interindividual corrrelates of lying confidence. The lines shows the relationship (with 95\\% CI uncertainty), assessed via Bayesian mixed models (\\***pd > 97\\%, \\***pd > 99\\%, \\****pd > 99.9\\%), between the participants' interoceptive and ToM composite scores and the confidence ratings of their responses. Average lying confidence (+/- 1 SD) within the two experimental conditions is displayed as points for descriptive purposes as the models were ran on individual trials."}
knitr::include_graphics("../figures/figure1.png")
```

## Interoception

The *Meta* interoception score was significantly associated with an increased confidence in lies, specifically in the polygraph condition ($\beta=0.20,~95\%~CI~[0.03, 0.35],~pd = 98.98\%$). It was also associated with faster answers for both lies ($\beta=-0.54,~95\%~CI~[-0.93, -0.15],~pd = 99.67\%$) and truths ($\beta=-0.29,~95\%~CI~[-0.63, 0.03],~pd = 96.10\%$), specifically in the polygraph condition. No significant association was found with heart rate.

The *Listening* interoception score was significantly associated with an increased confidence in lies, in both the polygraph ($\beta=0.43,~95\%~CI~[0.27, 0.59],~pd = 100\%$) and interrogation condition ($\beta=0.16,~95\%~CI~[0.01, 0.32],~pd = 98.04\%$). It was also associated with faster answers, particularly for lies ($\beta=-0.42,~95\%~CI~[-0.82, -0.03],~pd = 98.19\%$) and truths ($\beta=-0.36,~95\%~CI~[-0.76, 0.03],~pd = 96.49\%$) in the polygraph condition. No significant association was found with heart rate. This score also correlated with the dispositional lying *Contextuality* trait ($r = 0.50,~95\%~CI~[0.04, 0.64],~BF_{10} = 3.48\%$).

The *Focus* interoception score was significantly associated with an increased confidence in truthful responses in both the polygraph ($\beta=0.17,~95\%~CI~[-0.01, 0.34],~pd = 97.16\%$) and interrogation condition ($\beta=0.15,~95\%~CI~[-0.02, 0.32],~pd = 95.76\%$). No significant association was found with RT or heart rate, but the score correlated with the dispositional lying *Ability* trait ($r = 0.50,~95\%~CI~[0.22, 0.74],~BF_{10} = 34.37\%$).

The *Regulation* interoception score was significantly associated with an increased confidence in lies in both the polygraph ($\beta=0.32,~95\%~CI~[0.14, 0.51],~pd = 99.99\%$) and the interrogation conditions ($\beta=0.18,~95\%~CI~[0.00, 0.36],~pd = 97.42\%$), and with a decreased confidence in truth only in the polygraph condition ($\beta=-0.1,~95\%~CI~[-0.36, 0.01],~pd = 97.16\%$). No significant association was found with RT or heart rate (**Figure 2**).

```{r message=FALSE, warning=FALSE, out.width="100%", fig.cap="Summary of findings showing the positive (green) and negative (red) associations between interoception and theory of mind abilities and deception skills, depending on the experimental condition. It highlights that ToM was related to less confident and slower lies in the polygraph condition, and that specific interoceptive dimensions were related to more confident lies."}
knitr::include_graphics("../figures/figure2.png")
```

# Discussion

<!-- While there is an extensive body of literature associating deception with social cognition and emotion regulation, most research had been focused on their roles in deception detection, leaving a paucity in studies pertaining to their influence on lying ability.  -->

The present study examined the contribution of ToM and interoception on our ability to lie using a directed lie paradigm with two conditions ("Interrogation" and "Polygraph") designed to enhance each of the two mechanisms. Interestingly, we found that when participants were presented with (fake) physiological feedback (the polygraph condition), instead of a face of a person they had to lie to (the interrogation condition), their response time for both lies and truths increased, as did their heart rate. Although the condition did not impact the subjective confidence that participants had in their answers, it suggests that believing oneself to be submitted to a machine that is supposedly able to detect deception by interpreting physiological signals is a harder and/or more stressful condition than lying to a person. While research linking interoception and deception is limited, our results are in line with studies that show an association between interoceptive awareness and anxiety [@yoris2015roles; @garfinkel2013interoception; @domschke2010interoceptive]. Specifically, enhancing one's attention towards their internal bodily signals could have resulted in a hyper-vigilance towards physiological sensations that is perceived negatively. This is consistent with previous deception detection studies, in which participants' spontaneous lying behaviour only decreased when they were given feedback by a polygraph machine, but not when no feedback was given [@peleg2019lie]. By extension, our study adds to the controversial discourse surrounding the use of physiological measures in past deception research, further questioning its validity as an indicator of deception [@rosky2013f; @oviatt2018handbook].

<!-- This hence confers support for the manipulation used in our experimental paradigm, demonstrating that increasing the attentional focus towards visual interoceptive cues, even if they are fake, can elicit interoceptive awareness and enhance the precision of internal signals if such external cues of interoceptive signals are perceived to be real. Extending previous research pertaining to the application of interoceptive training for improving emotion regulation [@sugawara2020effects; @quadt2021interoceptive; @wu2022enhanced], our study thus provides evidence that interoceptive awareness can be manipulated and thereby modified to guide our emotional responses. -->

<!-- Interoceptive vs Social Cues -->
<!-- Using a bi-modal feedback deception task, we found partial support for our hypothesis. Specifically, participants reported lower lie confidence and took a longer time to respond, but were less physiologically aroused in the *Polygraph* condition, relative to the *Interrogation* condition. While this mixed finding is challenging to interpret, it is possible that one's interoceptive sensitivity, enhanced by the perceived interoceptive cues, interacted with the stress experienced in such deceptive interactions and led to a depletion of cognitive resources [@ten2015physically; @porter2012secrets]. Indeed, the experience of excessive stress has been shown to impair interoceptive accuracy, leading to a hyper-focus on internal bodily signals [@schulz2015interoception]. Furthermore, this is also in line with the established attentional switching hypothesis, which posits that an increase in attention towards internal signals and managing one's emotional reaction would confer less cognitive resources available for the actual act of telling a lie, thereby resulting in individuals taking a longer time to respond, and rating their lies with less confidence [@arnold2019interoception; @hanania2010selective]. This interplay of interoception and maladaptive stress response, and their effect on lying ability hence underscores the complexity underlying successful deceptive behaviour. -->

<!-- ToM -->
Our results suggest that higher ToM abilities were related to slower and less confident lies, but only in the polygraph condition. While previous bodies of work have reported mixed findings regarding the association between interoception and ToM [@scaffidi2016self; @gendolla2009self; @wundrack2023mindful; @chiou2013enactment], our results suggest the two are negatively linked. One possible interpretation of our findings is that people with stronger ToM abilities by default rely more on their social skills and altercentric inference when lying (i.e., they focus on - and try to read - the other person). When that mechanism is unavailable or unsuited (e.g., when there is no person to lie to - but a "machine" in our case), their corresponding lying ability decreases. However, in light of the current field of mixed findings relating interoception and ToM [@shah2017heart; @gao2019body; @canino2022embodiment; @miller2015connections], future studies are necessary to investigate the interaction of these mechanisms in different social contexts.

<!-- ionship between ToM and deception in contexts where interoceptive abilities are emphasized, the observed effects are in line with studies that suggest negative links associating ToM with interoceptive sensibility (i.e., the self-perceived tendency to monitor internal signals) [@canino2022embodiment].  -->

<!-- Interoception -->
We also showed that interoceptive abilities (as indicated by the composite interoception scores) are correlated with a higher confidence in one's lies in the polygraph condition, a condition in which the attention towards internal reactions is fostered. Indeed, this is in line with previous studies that found individuals with low interoception were more averse to risk when reputational stakes were high, telling fewer egoistical lies [@vabba2022interoceptive]. In fact, @vabba2022interoceptive further reports that people with high interoception abilities were less likely to differ in risk-taking tendencies, telling the same number of lies regardless of the social stakes. Consistent with our results, @von2023social further reports that individuals with high interoceptive accuracy were more likely to make egocentric decisions. However, in contrast to previous studies [@pinna2020systematic; @fustos2013embodiment; @owens2018investigating; @pollatos2007heart], we did not find any significant relationship between individuals' interoception scores and their heart rate changes during their answers. This points toward a predominantly meta-cognitive effect without necessarily an actual bodily regulation (i.e., participants with good interoception feel that their lies are more convincing, but do not actively attenuate their bodily reactions). 
<!-- [**also consistent with research demonstrating distinctions between interoceptive awareness and accuracy; extends pevious research in which individuals with greater interoceptive sensibility and accuracy were not better at detecting lies.**] -->

<!-- Our results show a significant association between individuals' composite ToM score with lower lie confidence and longer response time when interoceptive cues were highly salient. Under the free energy principle [@friston2010free], interoception is regarded as integral in ToM [@shah2017heart; @palmer2018going; @ondobaka2017role]. Specifically, during social inference, interoceptive signals are integrated with exteroceptive (e.g., visual and auditory) and proprioceptive information (i.e., motor) to guide one's social behaviour [@ondobaka2017role]. Given recent work in which ToM is associated with one having more precise interoceptive states [@tsakiris2017multisensory; @palmer2018going; @arnold2019interoception], it is possible that individuals with greater ToM abilities have higher interoceptive awareness, more stable self-other representations and less malleable percepts of self [@tsakiris2017multisensory; @arnold2019interoception]. As such, external cues of their interoceptive signals may be perceived as incongruent, hence generating a sense of uncertainty in their responses and leading to their lower lie confidence and longer response times.   -->

<!-- Meta-Interoception, Emotion Regulation and Lying -->
<!-- Consistent with our hypothesis, a significant positive association was predominantly found between metacognitive dimensions of interoception (*Meta* and *Regulation* interoception) and individuals' lie confidence when their physiological cues were salient. In contrast to previous studies [@pinna2020systematic; @fustos2013embodiment; @owens2018investigating; @pollatos2007heart], we did not find any significant relationship between individuals' interoception scores and objective heart rate. Taken together, these findings suggest that interoceptive awareness, in and of itself, is insufficient for adaptive emotion regulation. Indeed, some researchers have found positive links between interoceptive awareness and individuals' use of maladaptive strategies such as rumination [@tan2023interoceptive]. Furthermore, hypervigilance to bodily signals is also positively related to somatization and psychopathological disorders such as anxiety [@wang2020intrinsic; @dunn2010can; @pollatos2007interoceptive]. As such, further studies are necessary to examine what factors modulate the relationship between the metacognitve facets of interoception and emotion regulation. -->

<!-- Limitations and Future Research -->
Although yielding promising results, this exploratory study is limited by the sample size. Although we tried to mitigate it by 1) extracting more robust variables (by combining multiple ones by means of feature reduction) and 2) using a suited analysis approach (Bayesian statistics with informative priors), future replication studies with larger samples are warranted to confirm this first investigation. Nonetheless, we believe our results to be credible as we find consistent patterns across various facets and measures (for instance, all interoceptive dimensions, although distinct, share a similar trend) in line with theoretical expectations. The statistical power could also explain the overall lack of results found in relation to heart rate, which has a higher signal-to-noise ratio as compared to subjective reports (such as confidence scales).

Another aspect to note is the strong reliance on self-reported measures as outcome variables of lie ability (in particular, the measure of answer confidence, but also the auto-questionnaires). This might conflate meta-cognitive abilities as well as dishonest answers. Although we tried to include more objective measures, such as RT (although it too was tied to the participants' conscious decision to press a key) and heart rate, future studies should attempt at measuring objectively the answer (lie or truth) quality, for instance by means of external examiners. Note that this is not a limitation *per se*, as it answers a slightly different question - what are the correlates of *objective* lying skills - than deception self-confidence. 
<!-- [**look for lit linking to what are commonly considered cues for lying**] -->

Additionally, pertaining to the limitations with regards to the measure of lying ability, care has to be given to the measure of the predictor constructs, namely ToM and interoception. While we tried to include a behavioral task as well as a subjective questionnaire for each, it has to be underlined that they are notoriously difficult concepts to measure. In particular, objective interoceptive accuracy was assessed using the Heartbeat Counting Task (HCT). While the HCT used to be considered as a gold standard and remains one of the most commonly used measures [@desmedt2022does], concerns regarding its validity has been highlighted in several studies [@desmedt2018heartbeat; @desmedt2022does; @brener2016towards; @legrand2022heart]. Given increasing research efforts invested in the development of novel interoception tasks [@plans2021measuring; @legrand2022heart; @ponzo2021measuring], future works should further examine the relationship between interoception and lying ability using measures with better psychometric properties.

Moreover, although the cognitive and affective components of ToM and empathy share overlaps in the current literature, and there is no consensus regarding how the two concepts should be delineated, recent evidence nonetheless suggests ToM and empathy are necessarily distinct constructs with separable underlying mechanisms [@kanske2015dissecting]. As such, future studies are warranted to further investigate the associations between ToM and lying ability using validated instruments sensitive to measuring ToM (such as the Theory of Mind Inventory:Self-Report, @hutchins2021development). Furthermore, our application of feature reduction as a noise-elimination measure could have over-simplified the data. A more complex pattern of relationships, with different contributions of various subdimensions of ToM and interoception, could emerge provided a sufficient statistical power and valid measures.

<!-- Taking into perspective that the various facets of interoception are widely regarded as dissociable [@suksasilp2022towards; @mehling2012multidimensional], future studies could thus investigate their effects separately, to better delineate how interoceptive processes influence lying ability.    -->

In conclusion, this study is a first step towards assessing the contribution of ToM and interoception abilities in deception, particularly in one's ability to lie convincingly. To this end, we introduced a new paradigm that is able to modulate the contribution of these mechanisms while remaining relevant to applied fields of lie detection and criminology (in which the experimental conditions find echoing practices). Notably, our results provide princeps evidence that interoception could be a key - and overlooked - mechanism in deception. Furthermore, our findings extend and offer an alternate perspective to the debatable use of polygraphs, suggesting that its utility for lie detection is not only questionable, but could potentially enhance the deceptive ability of people with better interoception.

# Data Availability

The material (stimuli generation code, experiment code, raw data, analysis script with complementary figures and analyses, etc.) for this research is available at 
<!-- **https://github.com/DominiqueMakowski/DeceptionInteroTom**. -->
**[masked for blinding]**.

# Conflict of Interest Statement

The authors declare no conflict of interest.

# Acknowledgements

**[masked for blinding]**
<!-- We would like to thank Jia Rong Low and Maximilian Yong for their help in data collection. -->

\newpage

# References

```{r create_r-references}
r_refs(file = "references.bib")
```

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs}
:::

```{=tex}
\endgroup
```